### 2016-09-19 eg6-letters/eg6-letters/NIM/author_list.tex
merge files

\affiliation{Argonne National Laboratory, Argonne IL 60439, USA}
\affiliation{Institut de Physique Nucl$\acute{e}$aire, CNRS/IN2P3 and Universit$\acute{e}$ Paris Sud, Orsay, France}
\affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
%

\author{R.~Dupr\'e}   \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Institut de Physique Nucl$\acute{e}$aire, CNRS/IN2P3 and Universit$\acute{e}$ Paris Sud, Orsay, France}
\author{N.~Baltzell}  \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
\author{K.~Hafidi}    \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}
\author{M.~Hattawy}   \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Institut de Physique Nucl$\acute{e}$aire, CNRS/IN2P3 and Universit$\acute{e}$ Paris Sud, Orsay, France}
\author{S.~Stepanyan} \affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
%\author{} \affiliation{}
%

\collaboration{The JLab-CLAS Collaboration} 
%
\vskip 0.25cm