### 2016-12-22 analysis-note-vf/TitlePage.tex

\begin{titlepage}
\mainmatter
  {%\fontfamily{ptm}

       \vspace*{+3.0cm}

     \begin{center}
      \noindent {\Large \textbf{Analysis note}} \\

          \vspace*{+1.5cm}         \noindent {\Huge \textbf{Deeply virtual 
          Compton scattering}} \\
       \vspace*{0.15cm}
       \noindent {\Huge \textbf{off $^{4}He$}} \\
       \vspace*{1.0cm}
       \noindent \fontsize{16}{19}\selectfont (EG6 run)  \\
       \vspace*{0.8cm}

       \noindent \fontsize{14}{19}\selectfont M. Hattawy$^{[1]}$, N. Baltzell$^{[2,3]}$, R. Dupr\'e $^{[1]}$, H. Egiyan$^{[2]}$, L. El Fassi$^{[3,4]}$,\\ F.-X. Girod$^{[2]}$, K. Hafidi$^{[3]}$, C. Moody$^{[3]}$, S. Stepanyan$^{[2]}$  \\

       \vspace{+0.8cm}
        \noindent {\Large \textbf{Review committee:}} \\ 
      \vspace{+0.2cm}
       Michel Gar\c{c}on$^{[5]}$ (chair)\\
       Sebastian Kuhn$^{[6]}$ \\
       Zein-Eddine Meziani$^{[7]}$

       \vspace*{1.1cm}

        \normalsize $^{[1]}$Institut de Physique Nucl{\'e}aire d'Orsay, Orsay 91406, France \\
        \normalsize $^{[2]}$Jefferson Laboratory, Newport News, VA 230606, USA\\
        \normalsize $^{[3]}$Argonne National Laboratory, Argonne, IL 60439, USA \\
        \normalsize $^{[4]}$Mississippi State Unvirsity, Mississippi State, MS 39762, USA\\  
        \normalsize $^{[5]}$CEA-Saclay, Gif sur Yvette 91191, France\\  
        \normalsize $^{[6]}$Old Dominion University, Norfolk, VA 23529, USA\\
        \normalsize $^{[7]}$Temple University, Philadelphia, PA 19122, USA\\    
        \vspace{3.0cm}

    \today
     \end{center}

  }

  \newpage
  \thispagestyle{plain}

\end{titlepage}
\sloppy

\titlepage