### 2017-02-16 for_eg6_group/DVCS_EG6_meeting.tex
udate some studies for the incoherent channel

\section*{\centering{Tuning the incoherent exclusivity cuts}}

\begin{figure}[h!]
\includegraphics[height=15.0cm]{incoh/old_all_incoh_exc_cuts.pdf}
\caption{Old set of cuts}
\label{fig:cuts}
\end{figure}

\begin{figure}[h!]
   \includegraphics[height=15.0cm]{incoh/all_incoh_exc_cuts_test.pdf}
\caption{Suggested new set of cuts}
\label{fig:cuts}
\end{figure}

\begin{figure}[h!]
   \includegraphics[height=6.0cm]{incoh/ALU_phi_p_Q2.pdf}
   \includegraphics[height=5.2cm]{incoh/BSA_InCoherent_Q2.png}
\caption{}
\label{fig:cuts}
\end{figure}

\begin{figure}[h!]
   \includegraphics[height=6.0cm]{incoh/ALU_phi_p_x.pdf}
   \includegraphics[height=5.2cm]{incoh/BSA_InCoherent_xB.png}
\caption{}
\label{fig:cuts}
\end{figure}

\begin{figure}[h!]
   \includegraphics[height=6.0cm]{incoh/ALU_phi_p_t.pdf}
   \includegraphics[height=5.2cm]{incoh/BSA_InCoherent_t.png}
\caption{}
\label{fig:cuts}
\end{figure}