### 2017-02-20 for_eg6_group/incoh-2/Incoh.tex
add subdirectory for the fermi motion study

%\documentclass[a4paper,11pt,twoside]{ThesisStyle}
\documentclass[a4paper,11pt,twoside]{article}

\date{\today}
\include{formatAndDefs}
\begin{document}

\title{\centering{Investigating the Fermi motion effects on the incoherent DVCS 
channel}}
\maketitle

In our DVCS analysis, we considered the initial proton of the incoherent 
channel to be at rest, while this is not totally true because of the Fermi 
motion. Does this assumption is a good approximation? what other options can be 
used to minimize this effect on the calculated exclusive quantities?\\

Figure \ref{fig:handbag} presents the leading order handbag diagram of the 
incoherent DVCS channel off $^4$He, where the nucleus breaks and the DVCS 
occurs on a bound nucleon. Ideally, the transferred momentum squared ($t$) is 
defined using the initial and the final nucleons' 4-vectors as $(p-p')^{2}$.  
Then as a DVCS requirement, $t$ has to be greater than a certain value 
($t_{min}$), that is defined using the kinematics of the incoming and the 
scattered electrons. Due to the knowledge lack of the initial proton's 
momentum, this cut maybe inaccurate causes loosing some good DVCS events. In 
this work, we try to tackle this issue and check the stability of the 
reconstructed beam-spin asymmetry, as being our DVCS observable.\\

\begin{figure}[h!]
\includegraphics[height=10.0cm]{fig/handbag_incoherent.pdf}
\caption{Representation of the leading order handbag diagram of the incoherent 
DVCS process off $^4$He}
\label{fig:handbag}
\end{figure}

Typically, the transferred momentum squared $t$ is equal to the transferred 
momentum squared ($t'$) defined between the virtual photon and the final-state 
real photon, as shown in Figure \ref{fig:handbag}.  While defining the momentum 
transferred as $t'$ is a good way to pass the Fermi motion effect on the 
initial bound proton, it is more smeared than than $t$ due to bigger energy 
resolutions of the photons. Figure \ref{fig:tprime} shows the distributions of 
the coherent and the incoherent relative difference between $t$ and $t'$. One 
can see that the coherent DVCS channel, the distributions is symmetric and 
centered at zero as expected due to our good knowledge on the initial $^4$He 
target, while it is not exactly the case for the incoherent case due to the 
Fermi motion.\\

\begin{figure}[h!]
\includegraphics[height=5.0cm]{fig/T_tprime_Coh.png}
\includegraphics[height=5.0cm]{fig/T_tprime_InCoh.png}
\caption{On the left: the coherent relative difference between $t$ and $t'$.  
On the right: the incoherent relative difference.}
\label{fig:tprime}
\end{figure}

In the following, we check the effects of defining the transferred momentum 
squared as $t$ and $t'$ on the exclusive distributions and on the reconstructed 
beam-spin asymmetries. Figure \ref{fig:cuts_t} presents the exclusive 
distributions of the incoherent DVCS events where the transferred momentum
squared is defined as $t$, while it is define as $t'$ in figure 
\ref{fig:cuts_tprime}. To minimize the smearing effects of $t'$ on the 
exclusive distributions, 2$\sigma$ exclusive cuts were optimized compared to 
3$\sigma$ cuts which were applied in the $t$ case. In terms of collected 
statistics, the 2$\sigma$ cuts using $t'$ gave almost the same statistics using 
$t$. From comparing the exclusive distributions, especially the $ep\gamma$ 
missing energy and the missing mass squared, one sees that using $t'$ gives 
more Gaussian missing energy distributions and cleaner missing mass squared 
distribution.\\

\begin{figure}[h!]
\includegraphics[height=15.0cm]{fig/old_all_incoh_exc_cuts.pdf}
\caption{The old set of exclusive cuts with $t$>$t_{min}$. The black
   distributions represent the incoherent DVCS events candidate. The shaded
   distributions represent the events which passed all the exclusivity cuts
   except the quantity plotted. The vertical red lines represent the applied
exclusivity, 3$\sigma$ cuts.}
\label{fig:cuts_t}
\end{figure}

\begin{figure}[h!]
   \includegraphics[height=15.0cm]{fig/all_incoh_exc_cuts_test.pdf}
   \caption{New set of exclusive cuts with $t'$>$t_{min}$. The black
      distributions represent the incoherent DVCS events candidate. The shaded
      distributions represent the events which passed all the exclusivity cuts
      except the quantity plotted. The vertical red lines represent the applied 
   exclusivity, 2$\sigma$ cuts.}
\label{fig:cuts_tprime}
\end{figure}

Figure \ref{fig:bsa_incoh_bins} shows our measured incoherent beam-spin
asymmetries, for the three sets of the two-dimensional bins, in the two 
scinarios of defining the transferred momentum squared. The $Q^2$, $x_{B}$, and 
$-t$-dependencies of $A_{LU}$ at $\phi$=~90$^{\circ}$ (the $\alpha$ parameter 
of the fit) are shown in figure \ref{fig:incoh_Q2_xB_t_ALU}.

\begin{figure}[h!]
   \includegraphics[height=6.0cm]{fig/ALU_phi_p_Q2.pdf}
   \includegraphics[height=6.0cm]{fig/ALU_phi_p_x.pdf}
   \includegraphics[height=6.0cm]{fig/ALU_phi_p_t.pdf}
\caption{The reconstructed incoherent $A_{LU}$ as a function of the angle 
   $\phi$ in $Q^2$, $x_B$, and $-t$ bins, respectively from top to bottom using 
$t$>$t_{min}$ in black squares, and $t'$>$t_{min}$ in red squares.}
   \label{fig:bsa_incoh_bins}
\end{figure}

\begin{figure}[h!]
   \centering
   \includegraphics[height=6.0cm]{fig/ALU_90_p_vs_Q2_shortscenrario.pdf}\\
   \includegraphics[height=6.0cm]{fig/ALU_90_p_vs_x_shortscenrario.pdf}\\
   \includegraphics[height=6.0cm]{fig/ALU_90_p_vs_t_shortscenrario.pdf}
   \caption{The $Q^{2}$ (top), $x_{B}$ (middle), and $-t$-dependencies (bottom) 
      of the incoherent $A_{LU}$ at $\phi$~=~90$^{\circ}$, using the 
   $t$>$t_{min}$ cut and the corresponding exclusive cuts in black squares, and 
in red squares using the $t'$>$t_{min}$ and it's exclusive cuts.  The curves 
are theoretical calculations. On the bottom: the green circles are the HERMES 
$-A_{LU}$ (positron beam was used) inclusive measurements.}
\label{fig:incoh_Q2_xB_t_ALU}
\end{figure}

\section*{Conclusions}
In summary, we have studied the effects of Fermi motion on the measured 
incoherent beam-spin asymmetries by defining the transferred momentum squared 
through the nucleons in one case and through the photons in the other case. In 
terms of collected statistics, 2$\sigma$ exclusive cuts using the photons 
produced the same statistics as 3$\sigma$ cuts. The exclusive distributions 
using $t'$, through the photons, have shown more acceptable distributions in 
terms of being more symmetric in some cases and cleaner on others. The 
reconstructed asymmetries in the two cases of defining the transferred momentum 
squared are compatible and while the given error bars. As a suggestion, we 
would like to consider the $t'$ analysis as the final set for publications 
unless a major objections appear.

\end{document}