### 2017-06-08 eg6-letters/NIM/author_list.tex
Various fixes for NIM and coherent letter

\affiliation{Argonne National Laboratory, Argonne IL 60439, USA}
\affiliation{Institut de Physique Nucl$\acute{e}$aire, CNRS/IN2P3 and Universit$\acute{e}$ Paris Sud, Orsay, France}
\affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
%

\author{R.~Dupr\'e} \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Institut de Physique Nucl\'eaire, CNRS/IN2P3 and Universit\'e Paris Sud, Orsay, France}
\author{S.~Stepanyan} \affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
\author{M.~Hattawy} \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Institut de Physique Nucl\'eaire, CNRS/IN2P3 and Universit\'e Paris Sud, Orsay, France}
\author{N.~Baltzell} \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}\affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
\author{S.~Bueltmann} \affiliation{Old Dominion University, Norfolk, VA 23529, United States}
\author{A.~Celentano} \affiliation{INFN, Sezione di Genova, 16146 Genova, Italy}
\author{R.~De~Vita} \affiliation{INFN, Sezione di Genova, 16146 Genova, Italy}
\author{H.~Egiyan} \affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
\author{H.~Fenker} \affiliation{Jefferson Laboratory, Newport News, VA 230606, USA}
\author{K.~Hafidi} \affiliation{Argonne National Laboratory, Argonne IL 60439, USA}
\author{S.~Kuhn} \affiliation{Old Dominion University, Norfolk, VA 23529, United States}
\author{Y.~Perrin} \affiliation{LPSC, Universit\'e Grenoble-Alpes, CNRS/IN2P3, Grenoble, France}
\author{B.~Torayev} \affiliation{Old Dominion University, Norfolk, VA 23529, United States}
\author{E.~Voutier} \affiliation{Institut de Physique Nucl\'eaire, CNRS/IN2P3 and Universit\'e Paris Sud, Orsay, France}\affiliation{LPSC, Universit\'e Grenoble-Alpes, CNRS/IN2P3, Grenoble, France}
%\author{} \affiliation{}
%

\collaboration{The CLAS Collaboration} 
%
%\vskip 2.0cm