### 2017-06-12 eg6-letters/NIM/acknowledgment.tex
Add contact and acknowledgments.

\section{Acknowledgments}

The authors thank the staff of the Accelerator and Physics Divisions at the 
Thomas Jefferson National Accelerator Facility who made this work 
possible. This work was supported in part by the French Centre National de la 
Recherche Scientifique and the U.S. Department of Energy. One of the 
author (M.H.) also 
acknowledge the support of the Consulat G\'en\'eral de France \`a J\'erusalem. 
The Southeastern Universities Research Association operates the Thomas 
Jefferson National Accelerator Facility for the United States Department of 
Energy under contract DE-AC05-06OR23177.