### 2017-10-18 eg6-letters/NIM/acknowledgment.tex
more answers and implementations of the comments on the NIM paper

\section{Acknowledgments}

The authors thank the staff of the Accelerator and Physics Divisions at the 
Thomas Jefferson National Accelerator Facility who made this work possible.  
This work was supported in part by the French Centre National de la Recherche 
Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) 
and the U.S. Department of Energy. M.~Hattawy also acknowledges the support of 
the {\it Consulat G\'en\'eral de France \`a J\'erusalem}.  The Southeastern 
Universities Research Association operates the Thomas Jefferson National 
Accelerator Facility for the United States Department of Energy under contract 
DE-AC05-06OR23177. This material is based upon work supported by the U.S.  
Department of Energy, Office of Science, Office of Nuclear Physics, under 
contract number DE-AC02-06CH11357.