### 2018-04-25 eg6-letters/NIM/author_list.tex
Add e-mail of authors for communications

\author[anl,ipn]{R.~Dupr\'e\corref{cor}}\ead{dupre@ipno.in2p3.fr} 
\author[jlab]{S.~Stepanyan} 
\author[anl,ipn]{M.~Hattawy} 
\author[anl,jlab]{N.~Baltzell} 
\author[anl]{K.~Hafidi} 
\author[infn]{M.~Battaglieri} 
\author[odu]{S.~Bueltmann} 
\author[infn]{A.~Celentano} 
\author[infn]{R.~De~Vita} 
\author[anl,utfsm]{A.~El~Alaoui} 
\author[msu]{L.~El~Fassi} 
\author[jlab]{H.~Fenker} 
\author[anl]{K.~Kosheleva} 
\author[odu]{S.~Kuhn} 
\author[infn]{P.~Musico} 
\author[infn]{S.~Minutoli} 
\author[uc]{M.~Oliver} 
\author[lpsc]{Y.~Perrin} 
\author[odu]{B.~Torayev}
\author[ipn,lpsc]{E.~Voutier}

\address[anl]{Argonne National Laboratory, Argonne IL 60439, USA}
\address[ipn]{Institut de Physique Nucl\'eaire, CNRS/IN2P3 and Universit\'e Paris Sud, Orsay, France}
\address[jlab]{Jefferson Laboratory, Newport News, VA 230606, USA}
\address[infn]{INFN, Sezione di Genova, 16146 Genova, Italy}
\address[odu]{Old Dominion University, Norfolk, VA 23529, United States}
\address[utfsm]{Universidad T\'ecnica Federico Santa Mar\'ia, Casilla 110-V Valpara\'iso, Chile}
\address[msu]{Mississippi State University, Mississippi State, MS 39762-5167}
\address[uc]{University of Chicago, Chicago, IL 60637, United States}
\address[lpsc]{LPSC, Universit\'e Grenoble-Alpes, CNRS/IN2P3, Grenoble, France}

\cortext[cor]{Corresponding author}

%Author e-mail list: dupre@ipno.in2p3.fr, stepanya@jlab.org, mhattawy@anl.gov, baltzell@jlab.org, kawtar@anl.gov, marco.battaglieri@ge.infn.it, sbueltma@odu.edu, Andrea.Celentano@ge.infn.it, Raffaella.Devita@ge.infn.it, myahmed.elalaoui@usm.cl, elfassi@jlab.org, hcf@jlab.org, eakosheleva@gmail.com, skuhn@odu.edu, Paolo.Musico@ge.infn.it, saverio.minutoli@ge.infn.it, michael.oliver@yale.edu, perrin@lpsc.in2p3.fr, bayramtore@gmail.com, voutier@ipno.in2p3.fr