### 2018-07-05 eg6-letters/incoherent/incoherent_prl.tex
Add Stepan's comments

\RequirePackage{lineno}
\documentclass[twocolumn,nofootinbib,showpacs,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%
\def\Dirac#1{#1\hskip-5pt/}
%
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
{\linenomath\csname old#1\endcsname}%
{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
\patchAmsMathEnvironmentForLineno{#1}%
\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\begin{document}
\linenumbers

\title{Deeply Virtual Compton Scattering measurement off bound proton}

\input author_list.tex

\date{\today}
\begin{abstract}
 We report the first beam spin asymmetry (BSA) measurements in incoherent 
   deeply virtual Compton scattering (DVCS) off $^4$He. The data were 
   accumulated using a 6 GeV longitudinally 
   polarized electron beam incident on a pressurized $^4$He gaseous target 
   placed in front of the CLAS spectrometer in Hall-B at the Thomas Jefferson 
   National Accelerator Facility. The azimuthal angular ($\phi$) dependence of 
   the BSA is studied in a wide range of virtual photon and scattered proton 
   kinematics. The Q$^2$, $x_B$, and $t$ dependencies of the BSA on the bound 
   proton have been compared with the one on the free proton. In the whole 
   kinematical region of our measurements, the BSA on the bound proton is 
   smaller by $20\%$ to $40\%$, indicating possible medium modification 
   of its partonic structure. This measurement opens a new avenue for 
   studying structure of the bound nucleon and explore the origin of the EMC 
   effect using the framework of the generalized parton distributions.

\end{abstract}
\pacs{Valid PACS appear here}

\maketitle

Electromagnetic probes have played a major role in advancing our knowledge 
about the structure of the nucleon. While lepton-nucleon elastic scattering 
measurements have 
taught us about the spatial charge and magnetization distributions 
\cite{Hofstadter:1955ae,Perdrisat:2006hj}, deep-inelastic scattering 
experiments have uncovered the partonic structure of the nucleon and 
the longitudinal momentum distributions of the 
constituents partons, i.e., quarks and gluons \cite{pdg}.  
With nuclear targets, deeply inelastic lepton scattering measurements 
have revealed the fact that the 
distribution of quarks in a nucleus is not a simple convolution of 
their distributions within nucleons, an observation known as the ``EMC 
effect''\cite{EMC_first} (for reviews on the topic, see  
\cite{Arneodo:1992wf,Geesaman:1995yd,Norton:2003cb,Hen:2016kwk}).

A wealth of information on the structure of hadrons lies in the correlations 
between the momentum and spatial degrees of freedom of the partons. These 
correlations are set to be revealed through deeply virtual Compton scattering 
(DVCS), i.e., the hard exclusive lepto-production of a real photon, which 
provides access to a three-dimensional imaging of partons within the 
generalized parton distributions (GPDs) framework 
\cite{Mueller:1998fv,Ji:1996ek,Ji:1996nm,Radyushkin:1996nd,Radyushkin:1997ki}.   
The free proton DVCS measurement has been the focus of a worldwide effort 
\cite{Stepanyan:2001sm,Airapetian:2001yk,Airapetian:2006zr,Chekanov:2003ya,Aktas:2005ty,Chen:2006na,Munoz 
Camacho:2006hx,Girod:2007aa,Mazouz:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}
involving several accelerator facilities such as Jefferson Lab, DESY and  
CERN. These measurements now enable the extractions of GPDs 
and a 3-D tomography of the free proton \cite{Guidal:2013rya,Dupre:2016mai}.  
New measurements of DVCS from the $^{4}$He nucleus are a critical 
step towards providing a similar 3-D pictures of the quark structure of the 
nucleus~\cite{Dupre:2015jha}. In the nuclear case however, two channels are available,
the coherent channel where the nucleus is left intact \cite{Airapetian:2009cga,
Hattawy:2017woc}, and the incoherent channel where the DVCS happens on a nucleon.
The latter, which is the focus of this letter, provides a unique access to the 
modification of the partonic structure of the bound nucleons 
\cite{simonetta_2,Guzey:2006xi,Guzey:2008fe}. The 
$^{4}$He nucleus is an ideal experimental target for this measurement as it is 
characterized by a strong binding energy, a relatively high nuclear 
core density, and a large EMC effect~\cite{JSeely}. Moreover, it remains simple 
enough that precise calculation of its structure 
can be performed, making this nucleus the perfect target for our investigation of the 
medium modifications of the nucleon's partonic structure.

In the 
Bjorken regime, i.e, large virtual photon four-momentum squared 
($Q^{2}=-q^2=-(k-k^\prime)^2$), and at small invariant momentum transfer ($t=(p-p^\prime)^2$), the 
DVCS scattering process can be factorized, leaving the non-perturbative structure 
of the nucleon to be parametrized in terms of four chirally-even GPDs: $H$, 
$E$, $\widetilde{H}$, and $\widetilde{E}$, representing the four helicity-spin 
combinations of the quark-nucleon states \cite{Freund_Collins,Ji_Osborne}.

\begin{figure}[tb]
\includegraphics[width=7.5cm]{figs/handbag_incoherent.pdf}
\caption{Representation of the leading-order, twist-2, handbag diagram of the 
   incoherent DVCS process off $^4$He, where the four-vectors of the electrons, 
   photons, and protons are denoted by $k/k^\prime$, $q/q^\prime$, and 
   $p/p^\prime$, respectively. $x+\xi$ is the nucleon’s longitudinal momentum 
   fraction carried by the struck quark, 2$\xi$ is the longitudinal momentum 
   fraction of the momentum transfer $\Delta$ ($= p - p^\prime$), and 
   $t$~($=\Delta^2$) is the squared momentum transfer between the initial and
   the final state nucleon.}
\label{fig:diags}
\end{figure}

In this Letter, we present the first exclusive measurement of the beam-spin 
asymmetry (BSA) in deeply virtual electroproduction of a real photon off a bound 
proton in $^{4}$He. FIG.~\ref{fig:diags} illustrates the 
leading-twist handbag diagram for this process. Experimentally, we measure the squared sum of the 
Bethe-Heitler (BH) and the DVCS amplitudes. The BH process, 
where the real photon is emitted by the incident or the scattered electron 
rather than the nucleon, dominates the cross section at our kinematics. The BSA arises from the interference 
of these two terms and is directly sensitive to the DVCS amplitude that contain 
the information on the GPDs. Using a longitudinally polarized electron beam 
(L) and an unpolarized target (U), the BSA is defined as:
\begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the photon electroproduction 
differential cross section for a positive (negative) beam helicity.

Following the cross section decomposition provided in \cite{Belitsky:2001ns}, 
the different components can be expressed in terms of Fourier coefficients 
associated with $\phi$-harmonics, where $\phi$ is the angle between the 
leptonic and the hadronic planes of the reaction. At leading-twist, the 
BSA can be simplified as: 
\begin{equation}
   A_{LU}(\phi) = \frac{a_{0}\sin(\phi)}{1+a_{1}\cos(\phi)+a_{2}\cos(2\phi)}
   \label{eq:alu-simp}
\end{equation}
where the parameters $a_{0,1,2}$ are combinations of the above mentioned Fourier 
coefficients. The $\sin(\phi)$ harmonic is dominant in $A_{LU}$ 
and is proportional to the following combination of Compton form factors (CFF) 
\cite{Guidal:2013rya}
\begin{equation}
   a_{0} \propto \operatorname{Im}( F_1 \mathcal{H}- \frac{t}{4M^2} F_2 
   \mathcal{E}+ \frac{x_B}{2}(F_1+F_2)\tilde{\mathcal{H}})
\end{equation}
where $F_1$ and $F_2$ are the Dirac and Pauli form factors and $x_B$ the 
Bjorken scaling variable. The real and the imaginary parts of the CFF 
$\mathcal{H}$ relate to the GPD $H$ as  
\begin{align}
   \Re(&\mathcal{H}) = \mathcal{P} \int_{0}^{1}dx[H(x,\xi,t)-H(-x,\xi,t)] \, 
   C^{+}(x,\xi) \\
   \Im(&\mathcal{H}) = - \pi [H(\xi,\xi,t)-H(-\xi,\xi,t)],
\end{align}
with $\mathcal{P}$ the Cauchy principal value integral and $C^{+}$ a 
coefficient function defined as $(1/(x-\xi) + 1/(x+\xi))$, there $\xi$ is the 
skewing factor and can be
related to $x_B$ by $\xi\approx {{x_B}\over{2-x_B}}$. Similar expressions apply 
for the GPDs $E$, $\widetilde{H}$, and $\widetilde{E}$.

%experimental setup

The experiment (E08-024~\cite{Hafidi:2008pr}) took place in Hall-B of Jefferson 
Lab using the nearly 100\% duty factor, longitudinally-polarized electron beam 
(83$\%$ polarization) from the Continuous Electron Beam Accelerator Facility 
(CEBAF) at an energy of 6.064 GeV. The data were accumulated over 
40 days using a 6-atm-pressure, 292-mm-long, and 6-mm-diameter gaseous $^4$He 
target centered at 64~cm upstream of the CEBAF Large Acceptance Spectrometer 
(CLAS) coordinate center. For DVCS experiments, the CLAS baseline design 
\cite{Mecking:2003zu} is supplemented with an inner calorimeter (IC) and a 
solenoid magnet. The IC extends the photon detection acceptance of CLAS down 
to a polar angle of 4$^{\circ}$. The 5-Tesla solenoid magnet in the center of 
which the target was located prevented the high-rate low-energy M{\o}ller 
electrons from reaching the CLAS drift chambers by guiding these electrons 
inside a tungsten shield placed around the beam line.

% DVCS selection
Incoherent DVCS events were selected by requiring an electron, a proton, and at 
least one photon in the final state. Electrons were identified as negatively 
charged tracks reconstructed in the drift chambers with an appropriate energy 
deposition in the electromagnetic calorimeter (EC) and a light yield above the 
threshold in the Cherenkov counters. The protons were identified from 
positively charged tracks based on their time-of-flight measured 
by the scintillator counters.  In addition, it was required that both the 
electron and the proton in the event originated from the same vertex at the 
target. The photons were detected in 
either the IC or the CLAS electromagnetic calorimeter. All the clusters within 
the IC fiducial region were assumed to be photons, while a timing cut is used 
to clean the EC photons from neutrons. Note that even though the DVCS reaction 
has only one real photon in the final state, events with more than one photon 
were not discarded at this stage. These extra photons are mostly soft photons from accidental 
coincidence which, as will be discussed below, the DVCS exclusivity cuts 
eliminate easily. In the following selection stage, the most energetic photon 
is considered as the DVCS photon candidate (see \cite{Hattawy:thesis} for 
additional details on the particles identification).

Further requirements were applied to clean the identified initial set of 
incoherent DVCS events from accidental and physics background events. First, 
events were selected with Q$^{2}$ greater the 1 GeV$^2$ to ensure the 
applicability of the DVCS handbag diagram, and with $\gamma^{*}p$ invariant 
mass (W$=(q+p)^2$) greater than 2 GeV to avoid the nucleon resonance region.  
The $t$, transferred momentum squared to the recoil proton assuming an 
on-shell proton at rest, was required to be greater than a minimum allowed 
value at given Q$^2$ and W ($t~>~t_{min}$).  Then, the exclusivity of the 
incoherent DVCS events was ensured by imposing constraints on the kinematical 
variables resulting from conserving the four-momentum in the reaction 
$ep\rightarrow e'p'\gamma$. These kinematical variables are: the coplanarity 
angle $\Delta\phi$ between the ($\gamma,\gamma^*$) and ($\gamma^*$,$p'$) 
planes, the missing energy, mass, and transverse momentum of the $e'\gamma$ and 
$e'p'\gamma$ systems, the missing mass squared of the $e'p'$ system, and the 
angle $\theta$ between the measured photon and the missing momentum of the 
$e'p'$ system. The experimental data for the most relevant exclusivity 
variables are shown in FIG.~\ref{fig:kin-cuts}. We also rejected events where a 
$\pi^0$ was identified by the invariant mass of two photons. At the end of
this selection process, about 30k events 
passed all the requirements; their kinematic distributions are shown in 
FIG.~\ref{fig:kin-coverage}.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/incoh_exc_cuts_final.pdf}
\caption{The distributions 
   from left to right and from top to bottoms are: $\Delta \phi$, missing 
   energy, missing masses squared and the cone angle ($\theta$) between the 
   measured and the calculated photons in the $e'p'$ final-state system.
   The incoherent DVCS exclusivity cuts are represented by the vertical 
   red-dashed lines. The black distributions 
   represent the incoherent DVCS events candidate before the exclusivity 
   cuts. The shaded distributions represent the incoherent DVCS events which 
   passed all these cuts except the quantity plotted.}
\label{fig:kin-cuts}

\end{figure}
\begin{figure}[h!]
\hspace{-0.45cm}
\includegraphics[width=9.0cm]{figs/Q2_xB_t_InCoh.pdf}
\caption{Incoherent DVCS event distributions for $Q^{2}$ as a function of 
$x_{B}$ (left) and $Q^{2}$ as a function of $-t$ (right) after the exclusivity 
cuts.}
\label{fig:kin-coverage}
\end{figure}

%background
The two main backgrounds that contribute to the event sample after the exclusivity cuts are  
due to accidental coincidences and exclusive $\pi^0$ production where one of the photons 
from the $\pi^0$ decay escapes the detection. The contribution from accidental 
events, i.e., $e'p'\gamma$ collections 
with particles originating from different electron scatterings,
was evaluated to be of 6.5\% by selecting events passing all our selection cuts but originating from 
different vertices. The $\pi^0$ contamination was estimated using detector simulation and experimental 
data. From simulation, we calculated the ratio ($R = N^{1\gamma}_{sim}/N^{2\gamma}_{sim}$) 
of the number of $\pi^0$ 
events that are wrongly identified as exclusive $ep\rightarrow e'p'\gamma$ events to the 
number of events correctly identified as exclusive $ep\rightarrow e'p'\pi^0$. Then 
in each kinematical bin and for 
each beam-helicity state, the $\pi^0$-subtracted experimental DVCS events is 
calculated as $N = N^{ep\rightarrow e'p'\gamma}_{exp}- R~N^{ep\rightarrow 
e'p'\pi^0}_{exp}$, where $N^{ep\rightarrow e'p'\gamma}_{exp}$ ($N^{ep\rightarrow e'p'\pi^0}_{exp}$) is the 
number of $ep\rightarrow e'p'\gamma$ ($ep\rightarrow e'p'\pi^0$) experimentally identified. Depending on the 
kinematics, we found this contamination to be of 8 to 10\%.

%asymmetries
Experimentally, $A_{LU}$ is defined as
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} }.
\end{equation}
where $N^{+}$ and $N^{-}$ are the number of DVCS events for positive and negative 
beam-helicity state, and $P_{B}$ is the longitudinal beam polarization.

In the kinematical phase-space of our experiment, the GPD $H$ is
expected to dominate the 
contribution in the BSA over the other three GPDs. There, 
the $\phi$ dependence of $A_{LU}$ is mostly sensitive to the  
imaginary part of the CFF $\mathcal{H}$ through the parameters $a_0$ of 
equation~\ref{eq:alu-simp}. The measurement of the free proton 
$A_{LU}$ \cite{Girod:2007aa} has confirmed that the $a_0$ term is 
dominating the shape of the distribution.

Due to limited statistics, the data were binned two-dimensionally into 36 bins.  
That is, four bins in one of the kinematical variable of interest ($Q^{2}$, 
$x_{B}$ or $t$) and nine bins in the azimuthal angle ($\phi$).  
FIG.~\ref{fig:alu} presents the measured incoherent $A_{LU}$ as a function of 
$\phi$ in bins of $Q^{2}$, $x_{B}$, and $t$. The curves on the plots are fits 
of the form $\frac{a_{0}~sin(\phi)}{1+ a_{1}~cos(\phi)}$. The main contributions 
to systematic uncertainties on these fits are from the choice of the DVCS 
exclusivity cuts (6\%) and the large bin size (7\%). The systematic uncertainties
sum up to less than 10\% for all data points and thus always remain significantly 
smaller than the statistical errors.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/incoherent_ALU_phi.pdf}
\caption{The incoherent $A_{LU}$ as a function of $\phi$. Results are presented
   for different $Q^{2}$ bins (top panel), $x_{B}$ bins (middle panel), and 
   $-t$ bins (bottom panel). The error bars represent the statistical 
   uncertainties. The gray bands represent the systematic uncertainties, 
   including the normalization uncertainties. The black curves are the results 
   of our fits with the form $\frac{a_{0}~sin(\phi)}{1+ a_{1}~cos(\phi)}$.}
\label{fig:alu}
\end{figure}

We present in FIG.~\ref{fig:alu90} the dependence of the fitted $A_{LU}$ 
values at $\phi$~=~90$^{\circ}$ on the kinematical variables $Q^2$, $x_{B}$, 
and $t$. Within the given uncertainties, $A_{LU}$ does not indicate a strong 
dependence on $Q^2$. The $x_{B}$ and $t$ dependencies are compared to 
theoretical calculations performed by S.~Liuti and K.~Taneja 
\cite{simonetta_2}. Their model uses nuclear spectral function and 
considers mainly off-shell effects. The calculations are carried out at slightly 
different kinematics than our data but still provide important guidance. The 
experimental results appear to have smaller asymmetries especially at low 
$x_{B}$ than the calculations. These differences may arise from nuclear 
effects which are not taken into account in the model, such as long-range 
interactions. On the graph for $-t$ dependence, we show previous measurements 
by the HERMES collaboration \cite{Airapetian:2009cga}, in which only electrons
and photons were measured. Due to the large experimental uncertainties of the HERMES 
points, the two measurements are completely compatible.

\begin{figure}[tb]
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_Q2_shortscenrario.pdf}
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_x_shortscenrario.pdf}
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_t_shortscenrario.pdf}
\caption{The $Q^{2}$ (top), $x_{B}$ (middle), and $t$ dependencies (bottom) of
   the fitted $A_{LU}$ at $\phi$~=~90$^{\circ}$ (black squares). The error bars 
   represent the statistical uncertainties, while the gray bands represent the 
   systematic uncertainties. On the middle plot: the curves are theoretical 
   calculations from \cite{simonetta_2}. On the bottom: the solid (empty) green 
   circles are the HERMES $-A_{LU}$ (positron beam was used) inclusive 
measurements for the incoherent (coherent) enriched region 
\cite{Airapetian:2009cga}, the curves represent theoretical calculations from 
\cite{simonetta_2}.  } \label{fig:alu90}
\end{figure}

One can use the nuclear DVCS to measure a ``generalized'' EMC effect in order 
to see if significant nuclear effects are also visible within the GPD framework. 
To explore this idea, we construct 
the ratio of $A_{LU}$ for bound protons to that on free proton target.  
FIG.~\ref{fig:incoh_EMC_ratio_ALU_proton} presents the 
BSA ratio based on interpolation of the free proton asymmetries from  
CLAS \cite{Girod:2007aa} as a function of the kinematical variable
$x_B$. The $A_{LU}$ ratios show a 20\%-40\% lower asymmetries for a bound proton compared 
to the free proton mirroring a form of EMC effect but with a surprising 
large effect, in particular at low $x_B$. The measurements disagree with the off-shell \cite{simonetta_2} 
and the on-shell calculations that use the medium modified GPDs as calculated from 
the quark-meson coupling model \cite{Guzey:2008fe}. In particular, the difference becomes
large at lower $x_B$, 
where the off-shell model predicts an enhancement of the ratio, where our experiment 
shows a sharp drop. This measurement shows that an important nuclear effect is
missing from the existing models in order to explain this strong quenching of
the BSA.

\begin{figure}[tb]
\centering
%\includegraphics[width=6.9cm]{figs/ALU_ratioInc_Q2_shortscenrario.pdf}\\
\includegraphics[width=6.9cm]{figs/ALU_ratioInc_x_shortscenrario.pdf}
%\includegraphics[width=8.5cm]{figs/ALU_ratioInc_t_shortscenrario.pdf}
\caption{ The $A_{LU}$ ratio of the bound to the free proton at 
   $\phi$~=~90$^{\circ}$ as a function of $x_B$. The black squares are from this 
   work, the green circle is the HERMES measurement 
   \cite{Airapetian:2009cga}. The blue and red curves are results of off-shell 
   calculations \cite{simonetta_2}. The solid and dashed black curves are from 
   on-shell calculations \cite{Guzey:2008fe}.} 
   \label{fig:incoh_EMC_ratio_ALU_proton}
\end{figure}

%conclusion
In summary, we have presented our BSA measurements 
associated with incoherent DVCS off $^4$He using an upgraded setup of the CLAS 
spectrometer at Jefferson Lab. Our results were compared to model 
calculations based on different assumptions of the nuclear medium effects at 
the partonic level. The bound-proton BSA is largely suppressed compared to 
the free proton BSA. This surprising result opens a new avenue 
for progress in the understanding of the quarks and gluons behavior in the
nuclear medium.

%Acknowledgments

The authors acknowledge the staff of the Accelerator and Physics Divisions at 
the Thomas Jefferson National Accelerator Facility who made this experiment 
possible. This work was supported in part by the Chilean Comisi\'on Nacional de 
Investigaci\'on Cient\'ifica y Tecnol\'ogica (CONICYT), the Italian Instituto 
Nazionale di Fisica Nucleare, the French Centre National de la Recherche 
Scientifique, the French Commissariat \`a l'Energie Atomique, the U.S.  
Department of Energy under Contract No. DE-AC02-06CH11357, the United Kingdom 
Science and Technology Facilities Council (STFC), the Scottish Universities 
Physics Alliance (SUPA), the National Research Foundation of Korea, and the 
Office of Research and Economic Development at Mississippi State University.  
M.~Hattawy also acknowledges the support of the Consulat G\'en\'eral de France 
\`a J\'erusalem.  The Southeastern Universities Research Association operates 
the Thomas Jefferson National Accelerator Facility for the United States 
Department of Energy under Contract No. DE-AC05-06OR23177.

\begin{thebibliography}{99}

\bibitem{Hofstadter:1955ae} R.~Hofstadter and R.~W.~McAllister,
%``Electron Scattering From the Proton,''
Phys.\ Rev.\  {\bf 98}, 217 (1955).

\bibitem{Perdrisat:2006hj} C.~F.~Perdrisat, V.~Punjabi and M.~Vanderhaeghen,
%``Nucleon Electromagnetic Form Factors,''
Prog.\ Part.\ Nucl.\ Phys.\  {\bf 59}, 694 (2007)

\bibitem{pdg} J.~Beringer {\it et al.} (Particle Data Group), Phys.\ Rev.\ D 
   {\bf 86}, 010001 (2012).

\bibitem{EMC_first}
   J.~J.~Aubert {\it et al.}, 
      %“The ratio of the nucleon structure functions F 2 n for iron and
      %deuterium,” 
      Phys.\ Lett.\, vol.\ B { \bf 123}, pp. 275–278 (1983).

\bibitem{Arneodo:1992wf} 
  M.~Arneodo,
  %``Nuclear effects in structure functions,''
  Phys.\ Rept.\  {\bf 240}, 301 (1994).

\bibitem{Geesaman:1995yd} 
  D.~F.~Geesaman, K.~Saito and A.~W.~Thomas,
  %``The nuclear EMC effect,''
  Ann.\ Rev.\ Nucl.\ Part.\ Sci.\  {\bf 45}, 337 (1995).
  %doi:10.1146/annurev.ns.45.120195.002005

\bibitem{Norton:2003cb} 
  P.~R.~Norton,
  %``The EMC effect,''
  Rept.\ Prog.\ Phys.\  {\bf 66}, 1253 (2003).
  %doi:10.1088/0034-4885/66/8/201

\bibitem{Hen:2016kwk} 
  O.~Hen {\it et al.},
  %``Nucleon-Nucleon Correlations, Short-lived Excitations, and the Quarks Within,''
  Rev.\ Mod.\ Phys.\  {\bf 89}, no. 4, 045002 (2017).

\bibitem{Mueller:1998fv} D. Mueller, D. Robaschik, B. Geyer, F.M. Dittes, and 
   J.  Horejsi,
Fortsch.\ Phys. {\bf 42}, 101 (1994).

\bibitem{Ji:1996ek} 
X.D. Ji,
Phys.\ Rev.\ Lett. {\bf 78}, 610 (1997).

\bibitem{Ji:1996nm} 
X.D. Ji,
Phys.\ Rev.\ D {\bf 55}, 7114 (1997).

\bibitem{Radyushkin:1996nd}
A.V. Radyushkin,
Phys.\ Lett.\  B {\bf 380}, 417 (1996).

\bibitem{Radyushkin:1997ki} 
A.V. Radyushkin,
Phys.\ Rev.\ D {\bf 56}, 5524 (1997).

\bibitem{Stepanyan:2001sm}
S.~Stepanyan {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 87}, 182002 (2001).

\bibitem{Airapetian:2001yk} A.~Airapetian {\it et al.} [HERMES Collaboration],
%``Measurement of the beam spin azimuthal asymmetry associated with deeply 
   %virtual Compton scattering,''
 Phys.\ Rev.\ Lett.\  {\bf 87}, 182001 (2001)
%doi:10.1103/PhysRevLett.87.182001
%[hep-ex/0106068].

%\cite{Airapetian:2006zr}
\bibitem{Airapetian:2006zr} A.~Airapetian {\it et al.} [HERMES Collaboration],
 %``The Beam-charge azimuthal asymmetry and deeply virtual compton 
   %scattering,''
Phys.\ Rev.\ D {\bf 75}, 011103 (2007)
%doi:10.1103/PhysRevD.75.011103
%[hep-ex/0605108].

\bibitem{Chekanov:2003ya}
S. Chekanov {\it et al.} [ZEUS Collaboration],
Phys.\ Lett.\  B {\bf 573}, 46 (2003).

\bibitem{Aktas:2005ty}
A. Aktas {\it et al.} [H1 Collaboration],
Eur.\ Phys.\ J.\ C {\bf 44}, 1 (2005).

\bibitem{Chen:2006na} 
S.~Chen {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett.\ {\bf 97}, 072002 (2006).

\bibitem{Munoz Camacho:2006hx} 
C. Mu\~noz Camacho {\it et al.} [Jefferson Lab Hall A Collaboration],
Phys.\ Rev.\ Lett. {\bf 97}, 262002 (2006).

\bibitem{Girod:2007aa} 
F.X. Girod {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 100}, 162002 (2008).

\bibitem{Mazouz:2007aa} 
   M.~Mazouz {\it et al.} [Jefferson Lab Hall A Collaboration],
   Phys.\ Rev.\ Lett.\  {\bf 99}, 242501 (2007)

\bibitem{Gavalian:2009} 
G. Gavalian {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ C {\bf 80}, 035206 (2009).

\bibitem{Seder:2015} 
E. Seder {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 114}, 032001 (2015).

\bibitem{Pisano:2015} 
S.~Pisano {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ D {\bf 91}, 052014 (2015).

\bibitem{Jo:2015ema} H.~S.~Jo {\it et al.} [CLAS Collaboration],
  Phys.\ Rev.\ Lett.\  {\bf 115}, no. 21, 212003 (2015)

\bibitem{Guidal:2013rya} M.~Guidal, H.~Moutarde, and M.~Vanderhaeghen,
Rep.\ Prog.\ Phys.\  {\bf 76}, 066202 (2013).

\bibitem{Dupre:2016mai} R.~Dupre, M.~Guidal and M.~Vanderhaeghen,
 %``Tomographic image of the proton,''
 Phys.\ Rev.\ D {\bf 95}, no. 1, 011501 (2017)

\bibitem{Dupre:2015jha} 
  R.~Dupré and S.~Scopetta,
  Eur.\ Phys.\ J.\ A {\bf 52}, no. 6, 159 (2016)

%\cite{Airapetian:2009cga}
\bibitem{Airapetian:2009cga} A.~Airapetian {\it et al.} [HERMES Collaboration],
     %``Nuclear-mass dependence of azimuthal beam-helicity and beam-charge 
   %asymmetries in deeply virtual Compton scattering,''
Phys.\ Rev.\ C {\bf 81}, 035202 (2010)
%doi:10.1103/PhysRevC.81.035202
%[arXiv:0911.0091 [hep-ex]].

\bibitem{Hattawy:2017woc} 
   M.~Hattawy {\it et al.} [CLAS Collaboration],
   Phys.\ Rev.\ Lett.\  {\bf 119}, no. 20, 202004 (2017)

\bibitem{simonetta_2}
S.~Liuti and K.~Taneja, Phys.\ Rev.\ C {\bf 72}, 032201 (2005)

\bibitem{Guzey:2006xi} V.~Guzey and T.~Teckentrup, Phys.\ Rev.\ D {\bf 74}, 
   054027 (2006)

\bibitem{Guzey:2008fe} V.~Guzey, A.~W.~Thomas and K.~Tsushima,
  Phys.\ Lett.\ B {\bf 673}, 9 (2009)

\bibitem{JSeely}
J. Seely {\it et al.}, Phys.\ Rev.\ Lett.\ {\bf 103}, 202301 (2009).

\bibitem{Freund_Collins}
J.C.~Collins and A.~Freund, Phys.\ Rev.\ D {\bf 59}, 074009 (1999).

\bibitem{Ji_Osborne}
   X.-D.~Ji and J.~Osborne, Phys.\ Rev.\ D {\bf 58}, 094018 (1998).

\bibitem{Belitsky:2001ns}
A.~V.~Belitsky, D.~Mueller and A.~Kirchner,
Nucl.\ Phys.\ B {\bf 629}, 323 (2002)

\bibitem{Hafidi:2008pr} K.~Hafidi {\it et al.},
   proposal PR-08-024 to JLab PAC33 (unpublished).

\bibitem{Mecking:2003zu} B.~A.~Mecking {\it et al.} [CLAS Collaboration],
   Nucl.\ Instrum.\ Meth.\ A {\bf 503}, 513 (2003).

\bibitem{Dupre:2017upj} 
  R.~Dupr\'e {\it et al.},
  Nucl.\ Instrum.\ Meth.\ A {\bf 898}, 90 (2018)

%\cite{Armstrong:2017wfw}
\bibitem{Armstrong:2017wfw} W.~Armstrong {\it et al.},
     %``Partonic Structure of Light Nuclei,''
     arXiv:1708.00888 [nucl-ex].
       %%CITATION = ARXIV:1708.00888;%%
         %2 citations counted in INSPIRE as of 13 Feb 2018
     %\cite{Armstrong:2017zqr}
%\bibitem{Armstrong:2017zqr} W.~Armstrong {\it et al.},
       %``Tagged EMC Measurements on Light Nuclei,''
       arXiv:1708.00891 [nucl-ex].
         %%CITATION = ARXIV:1708.00891;%%
           %1 citations counted in INSPIRE as of 13 Feb 2018
       %\cite{Armstrong:2017zcm}
%    \bibitem{Armstrong:2017zcm} W.~R.~Armstrong {\it et al.},
         %``Spectator-Tagged Deeply Virtual Compton Scattering on Light 
       %Nuclei,''
         arXiv:1708.00835 [nucl-ex].
           %%CITATION = ARXIV:1708.00835;%%

\end{thebibliography}

\end{document}