### 2018-08-13 eg6-letters/incoherent/incoherent_prl.tex
implement Stepan's comments

\RequirePackage{lineno}
\documentclass[twocolumn,nofootinbib,showpacs,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%
\def\Dirac#1{#1\hskip-5pt/}
%
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
{\linenomath\csname old#1\endcsname}%
{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
\patchAmsMathEnvironmentForLineno{#1}%
\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\begin{document}
\linenumbers

\title{Deeply Virtual Compton Scattering measurement off bound protons}

\input author_list.tex

\date{\today}
\begin{abstract}
 We report the first beam spin asymmetry (BSA) measurement of incoherent deeply 
   virtual Compton scattering (DVCS) off $^4$He. The data were accumulated 
   using a 6 GeV longitudinally polarized electron beam incident on a 
   pressurized $^4$He gaseous target placed in front of the CLAS spectrometer 
   in Hall-B at the Thomas Jefferson National Accelerator Facility. The 
   azimuthal angular ($\phi$) dependence of the BSA is studied in a wide range 
   of virtual photon and scattered proton kinematics. The Q$^2$, $x_B$, and $t$ 
   dependencies of the BSA on the bound proton are compared with the one on the 
   free proton. In the whole kinematical region of our measurements, the BSA on 
   the bound proton is smaller by $20\%$ to $40\%$, indicating possible medium 
   modification of its partonic structure. This measurement opens a new avenue 
   for studying structure of the bound nucleon and explore the origin of the 
   EMC effect using the framework of the generalized parton distributions.

\end{abstract}
\pacs{Valid PACS appear here}

\maketitle

Electromagnetic probes have played a major role in advancing our knowledge 
about the structure of the nucleon. While lepton-nucleon elastic scattering 
measurements have 
taught us about the spatial charge and magnetization distributions 
\cite{Hofstadter:1955ae,Perdrisat:2006hj}, deep-inelastic scattering 
experiments have uncovered the partonic structure of the nucleon and 
the longitudinal momentum distributions of the constituent partons, i.e., 
quarks and gluons \cite{pdg}.  With nuclear targets, deeply inelastic lepton 
scattering measurements have revealed that the distribution of quarks in a 
nucleus is not a simple convolution of their distributions within nucleons, an 
observation known as the ``EMC effect''\cite{EMC_first} (for reviews on the 
topic, see  \cite{Arneodo:1992wf,Geesaman:1995yd,Norton:2003cb,Hen:2016kwk}).

A wealth of information on the structure of hadrons lies in the correlations 
between the momentum and spatial degrees of freedom of the partons. These 
correlations can be revealed through deeply virtual Compton scattering (DVCS), 
i.e., the hard exclusive lepto-production of a real photon, which provides 
access to a three-dimensional imaging of partons within the generalized parton 
distributions (GPDs) framework 
\cite{Mueller:1998fv,Ji:1996ek,Ji:1996nm,Radyushkin:1996nd,Radyushkin:1997ki}.   
The free proton DVCS measurement has been the focus of a worldwide effort 
\cite{Stepanyan:2001sm,Airapetian:2001yk,Airapetian:2006zr,Chekanov:2003ya,Aktas:2005ty,Chen:2006na,Munoz 
Camacho:2006hx,Girod:2007aa,Mazouz:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}
involving several accelerator facilities such as Jefferson Lab, DESY and  
CERN. These measurements now enable the extractions of GPDs 
and a 3-D tomography of the free proton \cite{Guidal:2013rya,Dupre:2016mai}.  
New measurements of DVCS from the $^{4}$He nucleus are a critical 
step towards providing a similar 3-D pictures of the quark structure of the 
nucleus~\cite{Dupre:2015jha}. In the nuclear case however, two channels are available,
the coherent channel where the scattering is off the entire nucleus, which is 
left intact in the final state \cite{Airapetian:2009cga,
Hattawy:2017woc}, and the incoherent channel where the DVCS happens on a 
nucleon, which is ejected from the nucleus.
The latter, which is the focus of this letter, provides a unique access to the 
modification of the partonic structure of the bound nucleons 
\cite{simonetta_2,Guzey:2006xi,Guzey:2008fe}. The 
$^{4}$He nucleus is an ideal experimental target for this measurement as it is 
characterized by a strong binding energy, a relatively high nuclear 
core density, and a large EMC effect~\cite{JSeely}. Moreover, it remains simple 
enough that precise calculation of its structure 
can be performed, making this nucleus the perfect target for our investigation of the 
medium modifications of the nucleon's partonic structure.

\begin{figure}[tb]
\includegraphics[width=7.5cm]{figs/handbag_incoherent.pdf}
\caption{Representation of the leading-order, twist-2, handbag diagram of the 
   incoherent DVCS process off $^4$He, where the four-vectors of the electrons, 
   photons, and protons are denoted by $k/k^\prime$, $q/q^\prime$, and 
   $p/p^\prime$, respectively. $x+\xi$ is the nucleon’s longitudinal momentum 
   fraction carried by the struck quark, 2$\xi$ is the longitudinal momentum 
   fraction of the momentum transfer $\Delta$ ($= p - p^\prime$), and 
   $t$~($=\Delta^2$) is the squared momentum transfer between the initial and
   the final state nucleon.}
\label{fig:diags}
\end{figure}

In this Letter, we present the first exclusive measurement of the beam-spin 
asymmetry (BSA) in deeply virtual electroproduction of a real photon off a bound 
proton in $^{4}$He. FIG.~\ref{fig:diags} illustrates the leading-twist handbag 
diagram for this process. In the Bjorken regime, i.e, large virtual photon 
four-momentum squared ($Q^{2}=-q^2=-(k-k^\prime)^2$), and at small invariant 
momentum transfer ($t=(p-p^\prime)^2$), the DVCS scattering process can be 
factorized, leaving the non-perturbative structure of the nucleon to be 
parametrized in terms of four chirally-even GPDs: $H$, $E$, $\widetilde{H}$, 
and $\widetilde{E}$, representing the four helicity-spin combinations of the 
quark-nucleon states \cite{Freund_Collins,Ji_Osborne}. Experimentally, we 
measure the squared sum of the Bethe-Heitler (BH) and the DVCS amplitudes. The 
BH process, where the real photon is emitted by the incident or the scattered 
electron rather than the nucleon, dominates the cross section at our 
kinematics. The BSA arises from the interference of these two terms and is 
directly sensitive to the DVCS amplitude that contains the information on the 
GPDs. Using a longitudinally polarized electron beam (L) and an unpolarized 
target (U), the BSA is defined as:
\begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the photon electroproduction 
differential cross section for a positive (negative) beam helicity.

Following the cross section decomposition provided in \cite{Belitsky:2001ns}, 
the different components can be expressed in terms of Fourier coefficients 
associated with $\phi$-harmonics, where $\phi$ is the angle between the 
leptonic and the hadronic planes of the reaction. At leading-twist, the 
BSA can be simplified as: 
\begin{equation}
   A_{LU}(\phi) = \frac{a_{0}\sin(\phi)}{1+a_{1}\cos(\phi)+a_{2}\cos(2\phi)}
   \label{eq:alu-simp}
\end{equation}
where the parameters $a_{0,1,2}$ are combinations of the above mentioned Fourier 
coefficients. The $\sin(\phi)$ harmonic is dominant in $A_{LU}$ 
and is proportional to the following combination of Compton form factors (CFF) 
\cite{Guidal:2013rya}
\begin{equation}
   a_{0} \propto \operatorname{Im}( F_1 \mathcal{H}- \frac{t}{4M^2} F_2 
   \mathcal{E}+ \frac{x_B}{2}(F_1+F_2)\tilde{\mathcal{H}})
\end{equation}
where $F_1$ and $F_2$ are the Dirac and Pauli form factors and $x_B$ the 
Bjorken scaling variable. The real and the imaginary parts of the CFF 
$\mathcal{H}$ relate to the GPD $H$ as  
\begin{align}
   \Re(&\mathcal{H}) = \mathcal{P} \int_{0}^{1}dx[H(x,\xi,t)-H(-x,\xi,t)] \, 
   C^{+}(x,\xi) \\
   \Im(&\mathcal{H}) = - \pi [H(\xi,\xi,t)-H(-\xi,\xi,t)],
\end{align}
with $\mathcal{P}$ the Cauchy principal value integral and $C^{+}$ a 
coefficient function defined as $(1/(x-\xi) + 1/(x+\xi))$, there $\xi$ is the 
skewing factor and can be
related to $x_B$ by $\xi\approx {{x_B}\over{2-x_B}}$. Similar expressions apply 
for the GPDs $E$, $\widetilde{H}$, and $\widetilde{E}$.

%experimental setup

The experiment (E08-024~\cite{Hafidi:2008pr}) took place in Hall-B of Jefferson 
Lab using the nearly 100\% duty factor, longitudinally-polarized electron beam 
(83$\%$ polarization) from the Continuous Electron Beam Accelerator Facility 
(CEBAF) at an energy of 6.064 GeV. The data were accumulated over 
40 days using a 6-atm-pressure, 292-mm-long, and 6-mm-diameter gaseous $^4$He 
target centered at 64~cm upstream of the CEBAF Large Acceptance Spectrometer 
(CLAS) coordinate center. For DVCS experiments, the CLAS baseline design 
\cite{Mecking:2003zu} is supplemented with an inner calorimeter (IC) and a 
solenoid magnet. The IC extends the photon detection acceptance of CLAS down 
to a polar angle of 4$^{\circ}$. The 5-Tesla solenoid magnet in the center of 
which the target was located prevented the high-rate low-energy M{\o}ller 
electrons from reaching the CLAS drift chambers by guiding these electrons 
inside a tungsten shield placed around the beam line.

% DVCS selection
Incoherent DVCS events were selected by requiring an electron, a proton, and at 
least one photon in the final state. Electrons were identified as negatively 
charged tracks reconstructed in the drift chambers with an appropriate energy 
deposition in the electromagnetic calorimeter (EC) and a light yield above the 
threshold in the Cherenkov counters. The protons were identified from 
positively charged tracks based on their time-of-flight measured 
by the scintillator counters.  In addition, it was required that both the 
electron and the proton in the event originated from the same vertex at the 
target. The photons were detected in either the IC or the CLAS electromagnetic 
calorimeter. All the clusters within the IC fiducial region were assumed to be 
photons, while a timing cut is used to clean the EC photons from neutrons. Note 
that even though the DVCS reaction has only one real photon in the final state, 
events with more than one photon were not discarded at this stage. These extra 
photons are mostly soft photons from accidental coincidence which, as will be 
discussed below, the DVCS exclusivity cuts eliminate easily. In the following 
selection stage, the most energetic photon is considered as the DVCS photon 
candidate (see \cite{Hattawy:thesis} for additional details on the particles 
identification).

Further requirements were applied to clean the identified initial set of 
incoherent DVCS events from accidental and physics background events. First, 
events were selected with Q$^{2}$ greater the 1 GeV$^2$ to ensure the 
applicability of the DVCS handbag diagram, and with $\gamma^{*}p$ invariant 
mass (W$=(q+p)^2$) greater than 2 GeV to avoid the nucleon resonance region. 
The $t$, transferred momentum squared to the recoil proton assuming an on-shell 
proton at rest, was required to be greater than a minimum allowed value at 
given Q$^2$ and W ($t>~t_{min}(=~- Q^{2}~\frac{2(1- x_{B})(1~-~\sqrt{1 
+~\epsilon^{2}})~+~\epsilon~^{2}}{4x_{B}(1-x_{B})+\epsilon^{2}}$,
where $\epsilon^{2}=~\frac{4M^{2}_{p}x^{2}_{B}}{Q^{2}}$ and $M_{p}$ is the 
proton mass). Then, the exclusivity of the incoherent DVCS events was ensured 
by imposing constraints on the kinematical variables resulting from conserving 
the four-momentum in the reaction $ep\rightarrow e'p'\gamma$.  These 
kinematical variables are: the coplanarity angle $\Delta\phi$ between the 
($\gamma,\gamma^*$) and ($\gamma^*$,$p'$) planes, the missing energy, mass, and 
transverse momentum of the $e'\gamma$ and $e'p'\gamma$ systems, the missing 
mass squared of the $e'p'$ system, and the angle $\theta$ between the measured 
photon and the missing momentum of the $e'p'$ system. The experimental data for 
the most relevant exclusivity variables are shown in FIG.~\ref{fig:kin-cuts}.  
We also rejected events where a $\pi^0$ was identified by the invariant mass of 
two photons. At the end of
this selection process, about 30k events 
passed all the requirements; their kinematic distributions are shown in 
FIG.~\ref{fig:kin-coverage}.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/incoh_exc_cuts_final.pdf}
\caption{The distributions 
   from left to right and from top to bottoms are: $\Delta \phi$, missing 
   energy, missing masses squared and the cone angle ($\theta$) between the 
   measured and the calculated photons in the $e'p'$ final-state system.
   The incoherent DVCS exclusivity cuts are represented by the vertical 
   red-dashed lines. The black distributions 
   represent the incoherent DVCS events candidate before the exclusivity 
   cuts. The shaded distributions represent the incoherent DVCS events which 
   passed all these cuts except the quantity plotted.}
\label{fig:kin-cuts}

\end{figure}
\begin{figure}[h!]
\hspace{-0.45cm}
\includegraphics[width=9.0cm]{figs/Q2_xB_t_InCoh.pdf}
\caption{Incoherent DVCS event distributions for $Q^{2}$ as a function of 
$x_{B}$ (left) and $Q^{2}$ as a function of $-t$ (right) after the exclusivity 
cuts.}
\label{fig:kin-coverage}
\end{figure}

%background
The two main backgrounds that contribute to the event sample after the exclusivity cuts are  
due to accidental coincidences and exclusive $\pi^0$ production where one of the photons 
from the $\pi^0$ decay escapes the detection. The contribution from accidental 
events, i.e., $e'p'\gamma$ collections 
with particles originating from different electron scatterings,
was evaluated to be of 6.5\% by selecting events passing all our selection cuts but originating from 
different vertices. The $\pi^0$ contamination was estimated using detector 
simulation and experimental data. From simulation, we calculated the ratio ($R 
= N^{1\gamma}_{sim}/N^{2\gamma}_{sim}$) of the number of $\pi^0$ events that 
are wrongly identified as exclusive $ep\rightarrow e'p'\gamma$ events to the 
number of events correctly identified as exclusive $ep\rightarrow e'p'\pi^0$.  
Then in each kinematical bin and for each beam-helicity state, the 
$\pi^0$-subtracted experimental DVCS events is calculated as $N = 
N^{ep\rightarrow e'p'\gamma}_{exp}- R~N^{ep\rightarrow e'p'\pi^0}_{exp}$, where 
$N^{ep\rightarrow e'p'\gamma}_{exp}$ ($N^{ep\rightarrow e'p'\pi^0}_{exp}$) is 
the number of the experimentally identified $ep\rightarrow e'p'\gamma$ 
($ep\rightarrow e'p'\pi^0$) events. Depending on the kinematics, we found this 
contamination to be of 8 to 10\%.

%asymmetries
Experimentally, $A_{LU}$ is defined as
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} }.
\end{equation}
where $N^{+}$ and $N^{-}$ are the number of DVCS events for positive and negative 
beam-helicity state, and $P_{B}$ is the longitudinal beam polarization.

In the kinematical phase-space of our experiment, the GPD $H$ is
expected to dominate the 
contribution in the BSA over the other three GPDs. There, 
the $\phi$ dependence of $A_{LU}$ is mostly sensitive to the  
imaginary part of the CFF $\mathcal{H}$ through the parameters $a_0$ of 
equation~\ref{eq:alu-simp}. The measurement of the free proton 
$A_{LU}$ \cite{Girod:2007aa} has confirmed that the $a_0$ term is 
dominating the shape of the distribution.

Due to limited statistics, the data were binned two-dimensionally into 36 bins.  
That is, four bins in one of the kinematical variable of interest ($Q^{2}$, 
$x_{B}$, or $t$) and then nine bins in the azimuthal angle ($\phi$).   
FIG.~\ref{fig:alu} presents the measured incoherent $A_{LU}$ as a function of 
$\phi$ in bins of $Q^{2}$ (integarted over full $x_{B}$ and $t$ ranges), 
$x_{B}$ (integarted over full $Q^{2}$ and $t$ ranges), and $t$ (integarted over 
full $Q^{2}$ and $x_{B}$ ranges). The curves on the plots are fits of the form 
$\frac{a_{0}~sin(\phi)}{1+ a_{1}~cos(\phi)}$.  The main contributions to 
systematic uncertainties on these fits are from the choice of the DVCS 
exclusivity cuts (6\%) and the large bin size (7\%). The systematic 
uncertainties sum up to less than 10\% for all data points and thus always 
remain significantly smaller than the statistical errors.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/incoherent_ALU_phi.pdf}
\caption{The incoherent $A_{LU}$ as a function of $\phi$. Results are presented
   for different $Q^{2}$ bins (top panel), $x_{B}$ bins (middle panel), and 
   $-t$ bins (bottom panel). The error bars represent the statistical 
   uncertainties. The gray bands represent the systematic uncertainties, 
   including the normalization uncertainties. The black curves are the results 
   of our fits with the form $\frac{a_{0}~sin(\phi)}{1+ a_{1}~cos(\phi)}$.}
\label{fig:alu}
\end{figure}

FIG.~\ref{fig:alu90} presents the dependence of the fitted $A_{LU}$ values at 
$\phi$~=~90$^{\circ}$ ($a_{0}$ parameter from the individual fits in 
FIG.~\ref{fig:alu}) on the kinematical variables $Q^2$, $x_{B}$, and $t$.  
Within the given uncertainties, $A_{LU}$ does not indicate a strong dependence 
on $Q^2$.  The $x_{B}$ and $t$ dependencies are compared to theoretical 
calculations performed by S.~Liuti and K.~Taneja \cite{simonetta_2}. Their 
model uses a nuclear spectral function and considers mainly off-shell effects.  
The calculations are carried out at slightly different kinematics than our data 
but still provide important guidance. The experimental results appear to have 
smaller asymmetries especially at low $x_{B}$ than the calculations. These 
differences may arise from nuclear effects which are not taken into account in 
the model, such as long-range interactions.  On the graph for $-t$ dependence, 
we show previous measurements by the HERMES collaboration 
\cite{Airapetian:2009cga}, in which only electrons
and photons were measured. Due to the large experimental uncertainties of the HERMES 
points, the two measurements are completely compatible.

\begin{figure}[tb]
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_Q2_shortscenrario.pdf}
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_x_shortscenrario.pdf}
\includegraphics[width=6.9cm]{figs/ALU_90_p_vs_t_shortscenrario.pdf}
\caption{The $Q^{2}$ (top), $x_{B}$ (middle), and $t$ dependencies (bottom) of
   the fitted $A_{LU}$ at $\phi$~=~90$^{\circ}$ (black squares). The error bars 
   represent the statistical uncertainties, while the gray bands represent the 
   systematic uncertainties. On the middle plot: the curves are theoretical 
   calculations from \cite{simonetta_2}. On the bottom: the solid (empty) green 
   circles are the HERMES $-A_{LU}$ (positron beam was used) inclusive 
measurements for the incoherent (coherent) enriched region 
\cite{Airapetian:2009cga}, the curves represent theoretical calculations from 
\cite{simonetta_2}.  } \label{fig:alu90}
\end{figure}

One can use the nuclear DVCS to measure a ``generalized'' EMC effect in order 
to see if significant nuclear effects are also visible within the GPD 
framework.  To explore this idea, we construct the ratio of $A_{LU}$ for bound 
protons to that on a free proton target. 
FIG.~\ref{fig:incoh_EMC_ratio_ALU_proton} presents the BSA ratio based on 
interpolation of the free proton asymmetries from  CLAS \cite{Girod:2007aa} as 
a function of the kinematical variable $x_B$. The $A_{LU}$ ratios show 
20\%-40\% lower asymmetries for a bound proton compared to the free proton 
mirroring a form of EMC effect but with a surprising large modification, in 
particular at low $x_B$. The measurements disagree with the off-shell 
\cite{simonetta_2} and the on-shell calculations that use the medium modified 
GPDs as calculated from the quark-meson coupling model \cite{Guzey:2008fe}. In 
particular, the difference becomes large at lower $x_B$, where the off-shell 
model predicts an enhancement of the ratio, while our experiment shows no sign 
of an increase of ratios. This measurement shows that an important nuclear 
effect is missing from the existing models in order to explain this strong 
quenching of the BSA.

\begin{figure}[tb]
\centering
%\includegraphics[width=6.9cm]{figs/ALU_ratioInc_Q2_shortscenrario.pdf}\\
\includegraphics[width=6.9cm]{figs/ALU_ratioInc_x_shortscenrario.pdf}
%\includegraphics[width=8.5cm]{figs/ALU_ratioInc_t_shortscenrario.pdf}
\caption{ The $A_{LU}$ ratio of the bound to the free proton at 
   $\phi$~=~90$^{\circ}$ as a function of $x_B$. The black squares are from this 
   work, the green circle is the HERMES measurement 
   \cite{Airapetian:2009cga}. The blue and red curves are results of off-shell 
   calculations \cite{simonetta_2}. The solid and dashed black curves are from 
   on-shell calculations \cite{Guzey:2008fe}.} 
   \label{fig:incoh_EMC_ratio_ALU_proton}
\end{figure}

%conclusion
In summary, we have presented our BSA measurements 
associated with incoherent DVCS off $^4$He using an upgraded setup of the CLAS 
spectrometer at Jefferson Lab. Our results were compared to model 
calculations based on different assumptions of the nuclear medium effects at 
the partonic level. The bound-proton BSA is largely suppressed compared to 
the free proton BSA. This surprising result opens a new avenue for progress in 
the understanding quarks and gluons in the nuclear medium.

%Acknowledgments

The authors acknowledge the staff of the Accelerator and Physics Divisions at 
the Thomas Jefferson National Accelerator Facility who made this experiment 
possible. This work was supported in part by the Chilean Comisi\'on Nacional de 
Investigaci\'on Cient\'ifica y Tecnol\'ogica (CONICYT), by CONICYT PIA grant 
ACT1413, the Italian Instituto Nazionale di Fisica Nucleare, the French Centre 
National de la Recherche Scientifique, the French Commissariat \`a l'Energie 
Atomique, the U.S.  Department of Energy under Contract No. DE-AC02-06CH11357, 
the United Kingdom Science and Technology Facilities Council (STFC), the 
Scottish Universities Physics Alliance (SUPA), the National Research Foundation 
of Korea, and the Office of Research and Economic Development at Mississippi 
State University.  M.~Hattawy also acknowledges the support of the Consulat 
G\'en\'eral de France \`a J\'erusalem.  The Southeastern Universities Research 
Association operates the Thomas Jefferson National Accelerator Facility for the 
United States Department of Energy under Contract No. DE-AC05-06OR23177.

\begin{thebibliography}{99}

\bibitem{Hofstadter:1955ae} R.~Hofstadter and R.~W.~McAllister,
%``Electron Scattering From the Proton,''
Phys.\ Rev.\  {\bf 98}, 217 (1955).

\bibitem{Perdrisat:2006hj} C.~F.~Perdrisat, V.~Punjabi and M.~Vanderhaeghen,
%``Nucleon Electromagnetic Form Factors,''
Prog.\ Part.\ Nucl.\ Phys.\  {\bf 59}, 694 (2007)

\bibitem{pdg} J.~Beringer {\it et al.} (Particle Data Group), Phys.\ Rev.\ D 
   {\bf 86}, 010001 (2012).

\bibitem{EMC_first}
   J.~J.~Aubert {\it et al.}, 
      %“The ratio of the nucleon structure functions F 2 n for iron and
      %deuterium,” 
      Phys.\ Lett.\, vol.\ B { \bf 123}, pp. 275–278 (1983).

\bibitem{Arneodo:1992wf} 
  M.~Arneodo,
  %``Nuclear effects in structure functions,''
  Phys.\ Rept.\  {\bf 240}, 301 (1994).

\bibitem{Geesaman:1995yd} 
  D.~F.~Geesaman, K.~Saito and A.~W.~Thomas,
  %``The nuclear EMC effect,''
  Ann.\ Rev.\ Nucl.\ Part.\ Sci.\  {\bf 45}, 337 (1995).
  %doi:10.1146/annurev.ns.45.120195.002005

\bibitem{Norton:2003cb} 
  P.~R.~Norton,
  %``The EMC effect,''
  Rept.\ Prog.\ Phys.\  {\bf 66}, 1253 (2003).
  %doi:10.1088/0034-4885/66/8/201

\bibitem{Hen:2016kwk} 
  O.~Hen {\it et al.},
  %``Nucleon-Nucleon Correlations, Short-lived Excitations, and the Quarks Within,''
  Rev.\ Mod.\ Phys.\  {\bf 89}, no. 4, 045002 (2017).

\bibitem{Mueller:1998fv} D. Mueller, D. Robaschik, B. Geyer, F.M. Dittes, and 
   J.  Horejsi,
Fortsch.\ Phys. {\bf 42}, 101 (1994).

\bibitem{Ji:1996ek} 
X.D. Ji,
Phys.\ Rev.\ Lett. {\bf 78}, 610 (1997).

\bibitem{Ji:1996nm} 
X.D. Ji,
Phys.\ Rev.\ D {\bf 55}, 7114 (1997).

\bibitem{Radyushkin:1996nd}
A.V. Radyushkin,
Phys.\ Lett.\  B {\bf 380}, 417 (1996).

\bibitem{Radyushkin:1997ki} 
A.V. Radyushkin,
Phys.\ Rev.\ D {\bf 56}, 5524 (1997).

\bibitem{Stepanyan:2001sm}
S.~Stepanyan {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 87}, 182002 (2001).

\bibitem{Airapetian:2001yk} A.~Airapetian {\it et al.} [HERMES Collaboration],
%``Measurement of the beam spin azimuthal asymmetry associated with deeply 
   %virtual Compton scattering,''
 Phys.\ Rev.\ Lett.\  {\bf 87}, 182001 (2001)
%doi:10.1103/PhysRevLett.87.182001
%[hep-ex/0106068].

%\cite{Airapetian:2006zr}
\bibitem{Airapetian:2006zr} A.~Airapetian {\it et al.} [HERMES Collaboration],
 %``The Beam-charge azimuthal asymmetry and deeply virtual compton 
   %scattering,''
Phys.\ Rev.\ D {\bf 75}, 011103 (2007)
%doi:10.1103/PhysRevD.75.011103
%[hep-ex/0605108].

\bibitem{Chekanov:2003ya}
S. Chekanov {\it et al.} [ZEUS Collaboration],
Phys.\ Lett.\  B {\bf 573}, 46 (2003).

\bibitem{Aktas:2005ty}
A. Aktas {\it et al.} [H1 Collaboration],
Eur.\ Phys.\ J.\ C {\bf 44}, 1 (2005).

\bibitem{Chen:2006na} 
S.~Chen {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett.\ {\bf 97}, 072002 (2006).

\bibitem{Munoz Camacho:2006hx} 
C. Mu\~noz Camacho {\it et al.} [Jefferson Lab Hall A Collaboration],
Phys.\ Rev.\ Lett. {\bf 97}, 262002 (2006).

\bibitem{Girod:2007aa} 
F.X. Girod {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 100}, 162002 (2008).

\bibitem{Mazouz:2007aa} 
   M.~Mazouz {\it et al.} [Jefferson Lab Hall A Collaboration],
   Phys.\ Rev.\ Lett.\  {\bf 99}, 242501 (2007)

\bibitem{Gavalian:2009} 
G. Gavalian {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ C {\bf 80}, 035206 (2009).

\bibitem{Seder:2015} 
E. Seder {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ Lett. {\bf 114}, 032001 (2015).

\bibitem{Pisano:2015} 
S.~Pisano {\it et al.} [CLAS Collaboration],
Phys.\ Rev.\ D {\bf 91}, 052014 (2015).

\bibitem{Jo:2015ema} H.~S.~Jo {\it et al.} [CLAS Collaboration],
  Phys.\ Rev.\ Lett.\  {\bf 115}, no. 21, 212003 (2015)

\bibitem{Guidal:2013rya} M.~Guidal, H.~Moutarde, and M.~Vanderhaeghen,
Rep.\ Prog.\ Phys.\  {\bf 76}, 066202 (2013).

\bibitem{Dupre:2016mai} R.~Dupre, M.~Guidal and M.~Vanderhaeghen,
 %``Tomographic image of the proton,''
 Phys.\ Rev.\ D {\bf 95}, no. 1, 011501 (2017)

\bibitem{Dupre:2015jha} 
  R.~Dupré and S.~Scopetta,
  Eur.\ Phys.\ J.\ A {\bf 52}, no. 6, 159 (2016)

%\cite{Airapetian:2009cga}
\bibitem{Airapetian:2009cga} A.~Airapetian {\it et al.} [HERMES Collaboration],
     %``Nuclear-mass dependence of azimuthal beam-helicity and beam-charge 
   %asymmetries in deeply virtual Compton scattering,''
Phys.\ Rev.\ C {\bf 81}, 035202 (2010)
%doi:10.1103/PhysRevC.81.035202
%[arXiv:0911.0091 [hep-ex]].

\bibitem{Hattawy:2017woc} 
   M.~Hattawy {\it et al.} [CLAS Collaboration],
   Phys.\ Rev.\ Lett.\  {\bf 119}, no. 20, 202004 (2017)

\bibitem{simonetta_2}
S.~Liuti and K.~Taneja, Phys.\ Rev.\ C {\bf 72}, 032201 (2005)

\bibitem{Guzey:2006xi} V.~Guzey and T.~Teckentrup, Phys.\ Rev.\ D {\bf 74}, 
   054027 (2006)

\bibitem{Guzey:2008fe} V.~Guzey, A.~W.~Thomas and K.~Tsushima,
  Phys.\ Lett.\ B {\bf 673}, 9 (2009)

\bibitem{JSeely}
J. Seely {\it et al.}, Phys.\ Rev.\ Lett.\ {\bf 103}, 202301 (2009).

\bibitem{Freund_Collins}
J.C.~Collins and A.~Freund, Phys.\ Rev.\ D {\bf 59}, 074009 (1999).

\bibitem{Ji_Osborne}
   X.-D.~Ji and J.~Osborne, Phys.\ Rev.\ D {\bf 58}, 094018 (1998).

\bibitem{Belitsky:2001ns}
A.~V.~Belitsky, D.~Mueller and A.~Kirchner,
Nucl.\ Phys.\ B {\bf 629}, 323 (2002)

\bibitem{Hafidi:2008pr} K.~Hafidi {\it et al.},
   proposal PR-08-024 to JLab PAC33 (unpublished).

\bibitem{Mecking:2003zu} B.~A.~Mecking {\it et al.} [CLAS Collaboration],
   Nucl.\ Instrum.\ Meth.\ A {\bf 503}, 513 (2003).

\bibitem{Hattawy:thesis}
M.~Hattawy, Ph.D. thesis, Universit{\'e} Paris Sud - Paris XI, France, 2015 
[Institution Report No. 2015PA112161].

\bibitem{Dupre:2017upj} 
  R.~Dupr\'e {\it et al.},
  Nucl.\ Instrum.\ Meth.\ A {\bf 898}, 90 (2018)

%\cite{Armstrong:2017wfw}
\bibitem{Armstrong:2017wfw} W.~Armstrong {\it et al.},
     %``Partonic Structure of Light Nuclei,''
     arXiv:1708.00888 [nucl-ex].
       %%CITATION = ARXIV:1708.00888;%%
         %2 citations counted in INSPIRE as of 13 Feb 2018
     %\cite{Armstrong:2017zqr}
%\bibitem{Armstrong:2017zqr} W.~Armstrong {\it et al.},
       %``Tagged EMC Measurements on Light Nuclei,''
       arXiv:1708.00891 [nucl-ex].
         %%CITATION = ARXIV:1708.00891;%%
           %1 citations counted in INSPIRE as of 13 Feb 2018
       %\cite{Armstrong:2017zcm}
%    \bibitem{Armstrong:2017zcm} W.~R.~Armstrong {\it et al.},
         %``Spectator-Tagged Deeply Virtual Compton Scattering on Light 
       %Nuclei,''
         arXiv:1708.00835 [nucl-ex].
           %%CITATION = ARXIV:1708.00835;%%

\end{thebibliography}

\end{document}