### 2018-10-04 incoh_fermi_investigation/Incoh_fermi.tex
added documenation for fermi motion effect on incoherent dvcs

%\documentclass[a4paper,11pt,twoside]{ThesisStyle}
\documentclass[a4paper,11pt,twoside]{article}

\date{\today}
\include{formatAndDefs}
\begin{document}

\title{\centering{Investigating the Fermi motion smearing on the incoherent 
DVCS channel}}
\maketitle

\begin{figure}[h!]
\includegraphics[height=10.0cm]{figs/t_n_Incoh.pdf}
\caption{}
\label{fig:exp}
\end{figure}

\begin{figure}[h!]
\includegraphics[height=10.0cm]{figs/sim_Fermi_mom.png}
\caption{}
\label{fig:fermi}
\end{figure}

\begin{figure}[h!]
\includegraphics[height=10.0cm]{figs/sim_t_n_Incoh.pdf}
\caption{}
\label{fig:fermi}
\end{figure}

\end{document}