### 2018-12-21 eg6-letters/incoherent/publication_version/edit_author_list.tex
add version of the paper that fits the publication procedures

M. Hattawy, N.A. Baltzell, R. Dupr\'{e}, S. B\"{u}ltmann, R. De Vita, A. El Alaoui, L. El Fassi, H. Egiyan, F.X. Girod, M. Guidal, K. Hafidi, D. Jenkins, S. Liuti, Y. Perrin, S. Stepanyan, B. Torayev, E. Voutier, S. Adhikari, Giovanni Angelini, C. Ayerbe Gayoso, L. Barion, M. Battaglieri, I. Bedlinskiy, A.S. Biselli, F. Boss\`u, W. Brooks, F. Cao, D.S. Carman, A. Celentano, P. Chatagnon, T. Chetry, G. Ciullo, L. Clark, P.L. Cole, M. Contalbrigo, V. Crede, A. D'Angelo, N. Dashyan, E. De Sanctis, M. Defurne, A. Deur, S. Diehl, C. Djalali, M. Ehrhart, P. Eugenio, S. Fegan, A. Filippi, T.A.Forest, A. Fradi, M. Gar\c{c}on, G. Gavalian, N. Gevorgyan, G.P. Gilfoyle, K.L. Giovanetti, E. Golovatch, R.W. Gothe, K.A. Griffioen, N. Harrison, F. Hauenstein, T.B. Hayward, D. Heddle, K. Hicks, M. Holtrop, Y. Ilieva, D.G. Ireland, E.L. Isupov, S. Johnston, D. Keller, G. Khachatryan, M. Khachatryan, A. Khanal, M. Khandaker, C.W. Kim, W. Kim, F.J. Klein, V. Kubarovsky, S.E. Kuhn, L. Lanza, M.L. Kabir, P. Lenisa, K. Livingston, I.J.D. MacGregor, D. Marchand, N. Markov, M. Mayer, B. McKinnon, Z.E. Meziani, T. Mineeva, M. Mirazita, R.A. Montgomery, C. Munoz Camacho, P. Nadel-Turonski, S. Niccolai, A.I. Ostrovidov, L.L. Pappalardo, R. Paremuzyan, E. Pasyuk, O. Pogorelko, J. Poudel, Y. Prok, D. Protopopescu, M. Ripani, D. Riser, A. Rizzo, G. Rosner, P. Rossi, F. Sabati\'e, C. Salgado, R.A. Schumacher, Y.G. Sharabian, Iu. Skorodumina, D. Sokhan, O. Soto, N. Sparveris, S. Strauch, M. Taiuti, J.A. Tan, N. Tyler, M. Ungaro, H. Voskanyan, R. Wang, D.P. Watts, X. Wei, L.B. Weinstein, M.H. Wood, N. Zachariou, J. Zhang, Z.W. Zhao (The CLAS Collaboration)