### 2019-05-23 eg6-letters/incoherent/LZ15969_Hattawy_SummaryOfChanges_May23_2019.tex
add updated figure 4 and response to the referees

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{LZ15969:  Authors' Reply}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Summary of Changes}
Below is a summary list of the changes to the paper:

\begin{itemize}
   \item{ In response to Referee A (remove ``only''): it has been removed.}
   \item{ In response to Referee B (remove ``to ensure
the applicability of the DVCS handbag diagram''): it has been removed.}
   \item{ In response to Referee B (change ``avoid'' to ``avoids''): it has 
      been replaced.}
   \item { The figures has been edited to be within the maximum allowed length 
      3750.}

\end{itemize}

\end{document}