### 2019-05-24 eg6-letters/incoherent/LZ15969_Hattawy_SummaryOfChanges_May23_2019.tex
update reply

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{LZ15969:  Authors' Reply}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Summary of Changes}
We agree with the referees' suggestions and implemented the changes to the paper:

\begin{itemize}
   \item{ Per the recommendation of Referee A, we removed the ``only'' in the abstract.}
   \item{ Per the recommendation of Referee B, we removed ``to ensure the applicability of the DVCS handbag diagram'' and changed ``avoid'' to ``avoids'' at the end of page 3.}
   \item{ The figures have been adjusted to meet the maximum allowed length requirement.}   
\end{itemize}

\end{document}