### 2019-06-25 eg6-letters/incoherent/journal-proof/proof_cover2/authors_response_journal_proof2.tex
add the response to the second proof

%\documentclass[a4paper,11pt,twoside]{ThesisStyle}
\documentclass[a4paper,11pt,twoside]{article}

\title{Second Journal proof on Article LZ15969}

\date{\today}
\include{formatAndDefs}

\begin{document}

\maketitle

\begin{enumerate}

\item Second Proof: Please check and confirm the edit made in the sentence 
   ``With nuclear targets, deeply inelastic lepton scattering
...''\\
   \textcolor{blue}{Response: We are fine with adding ``European Muon 
      Collaboration (EMC) effect'' at the end of the sentence.}

\item  Second Proof: Please carefully confirm that all first-proof corrections 
   were addressed and that changes were made accurately. \\
   \textcolor{blue}{Response: Everything is well implemented.    }

\item Second Proof: Please confirm the paper is now ready to be published in 
   its current form.\\  \textcolor{blue}{Response: Everything looks fine except 
      reference 41. Right click on the link somehow does not open the link 
      right away. We confirm that the link 
      \url{https://tel.archives-ouvertes.fr/tel-01218756/document} is right.  }

\end{enumerate}

\end{document}