### 2019-07-09 eg6-letters/incoherent/journal-proof/proof_cover3/authors_response_journal_proof3.tex
Add the third proofs for the incoherent paper

%\documentclass[a4paper,11pt,twoside]{ThesisStyle}
\documentclass[a4paper,11pt,twoside]{article}

\title{Third Journal proof on Article LZ15969}

\date{\today}
\include{formatAndDefs}

\begin{document}

\maketitle

\begin{enumerate}

\item Third proof: Please check very closely to confirm all of the corrections have been incorporated properly throughout this
Letter and also approve the Letter is ready to be published in its current form.\\
   \textcolor{blue}{Response: See next.}

\item Third proof: Kindly check and confirm whether the URL is now working. \\
   \textcolor{blue}{Response: The link does not appear to be working.}

\end{enumerate}

\end{document}