### 2020-03-25 eg6-letters/long-paper/prc-dvcshelium.tex
Add corrections to long paper

\documentclass[aps,prc,preprint,superscriptaddress]{revtex4}
%\usepackage{amsmath,amssymb}             % AMS Math
%\usepackage[francais]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern,textcomp}
%\usepackage[left=3.5cm,right=2.5cm,top=2cm,bottom=2cm,includefoot,includehead,headheight=15pt]{geometry}
\usepackage{setspace}
\usepackage{microtype}
\usepackage{epigraph}
\usepackage{natbib}
%\usepackage[scaled]{helvet}
%\usepackage[pagebackref,hyperindex=true]{hyperref}
%\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{array}
\usepackage{amsmath}

\begin{document}

%Title of paper
\title{Measurement of deeply virtual Compton scattering off Helium-4 with CLAS at Jefferson Lab}

\input author_list.tex

\date{\today}

\begin{abstract}
Measurements of the beam-spin asymmetry in the deeply virtual Compton 
scattering off $^4$He using the CEBAF Large Acceptance Spectrometer (CLAS) are 
reported. A 6 GeV longitudinally polarized electron beam was incident on a 
pressurized $^4$He gaseous target to study the partonic structure of the 
nucleus and the bound nucleons. To ensure the exclusivity of the reactions, 
CLAS was upgraded with a radial time projection chamber to detect the 
low-energy recoiling $^4$He nuclei and an inner calorimeter to extend the 
photon detection acceptance at forward angles. Our results confirm the 
theoretically predicted enhancement of the coherent 
($e^4$He~$\to~e'$$^4$He$'\gamma'$) beam-spin asymmetries compared to those 
observed on the free proton, while the incoherent 
($e^4$He~$\to~e'$p$'\gamma'$X$'$) asymmetries exhibit a 30$\%$ suppression.  
From the coherent data, we were able to extract, in a model-independent way, 
the real and imaginary parts of the only $^4$He Compton form factor, $\cal 
H_A$, leading the way toward 3D imaging of the partonic structure of nuclei.
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{}

\maketitle

\section{Introduction}

In the past few decades, the study of the proton structure has made significant progress
thanks to the theoretical and experimental developments with three dimensional 
structure functions \cite{Anselmino:2015uka}. These studies, which have focused on generalized
parton distributions (GPD) and transverse momentum dependent parton distribution functions (PDF)
can be generalized to the nucleus and offer a unique opportunity to revisit the quark structure 
of the nucleus with an original perspective~\cite{Dupre:2015jha}. This new approach is 
particularly needed as the quark structure of the nucleus remains today the subject of numerous
controversies. Indeed, while much progress has been made in measuring the nuclear
PDFs, their shape remains fundamentally unexplained \cite{Norton:2003cb,Malace:2014uea,Hen:2016kwk}.

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions in the nuclear 
medium and offers a unique opportunity to make a measurement free of final state interactions 
(FSI). Moreover, with a nuclear targets of spin-0 the extraction of the GPD from the DVCS data 
is significantly simplified with the reduction to a single GPD intervening in the process at leading order. The 
measurement of the nuclear DVCS is however challenging experimentally, first attempts from 
the HERMES collaboration \cite{Airapetian:2009cga} lead to controversial conclusions. We present here
in detail the more recent measurements from the CLAS collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the nuclear DVCS CLAS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

\section{Theoretical Framework}

\subsection{The GPD Formalism}

The theory of GPDs has been already reviewed in details in various 
publications~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya},
we summarized here only necessary elements to discuss our experimental results.
The GPDs are real structure functions $F^{q}(x,\xi,t)$, where $x+\xi$ and $x-\xi$ are incoming and outgoing quark momenta 
respectively and $t=\Delta^2$ the squared transferred momentum to the target in this exchange of quarks as illustrated in 
Fig. \ref{fig:GPD}.

\begin{figure}[tbp!]
\center
\includegraphics[width=10.0cm]{fig1/GPD.png}
\caption{General representation for the GPDs of a nucleon represented by the triple lines and noted $N$.
	Single lines can represent quarks or anti-quarks probed in the nucleon.}
\label{fig:GPD}
\end{figure}

The different spin states possible lead to several independent GPDs for any 
given hadron. The proper accounting of the number of GPDs must be done with 
regard to the symmetries of the system. At leading order and leading twist, we 
find that there are $2(2J+1)^2$ GPDs for a particle of spin $J$, 
$i.e.$ for a spin-0 hadron, like the helium-4 nucleus, we will have two GPDs, 
for a spin-1/2 hadron, like the proton, eight GPDs. Half of these involve a 
parton helicity flip, they are called transversity GPDs and do not intervene 
in the DVCS process.

DVCS is the main experimental access to GPDs. However, this process does not offer
a full access to the GPDs and we only access them integrated over $x$, in the following form:
\begin{equation}
	\int^1 _{-1} dx F^{q}(x,\xi,t) \left [ {1 \over x - \xi - i \epsilon} 
	                                   \mp {1 \over x + \xi - i \epsilon} \right]. 
\end{equation}
To account for this and simplify equations, we define the complex Compton form factors (CFF,
noted with curved $\mathcal{F}$ for a given GPD $F$) for each GPD as follows:
\begin{equation}
	\Re e (\mathcal{F}(\xi,t)) = \sum _q e^2_q \mathcal{P} \int^1_{-1} dx F^{q}(x,\xi,t)
	      \left [ {1 \over x - \xi} \mp {1 \over x + \xi} \right], 
\end{equation}
\begin{equation}
	\Im m (\mathcal{F}(\xi,t)) = - \pi \sum _q e^2_q \left [ F^{q}(\xi,\xi,t) \mp F^{q}(-\xi,\xi,t) \right]. 
\end{equation} 
These are the quantities directly present in the DVCS cross sections. We note that they are summed over
the different quark flavors present in the hadron, as the electromagnetic probe do not differentiate
quark flavors.

Experimentally, another process is indistinguishable from the DVCS, the Bethe-Heitler (BH) 
process in which 
the final state photon is emitted by the scattering lepton rather than the hadron. In this case, the 
photon-hadron interaction is the same as in elastic scattering and depends on the
target FFs rather than its GPDs. The DVCS and BH processes are experimentally indistinguishable as they
have identical final states, such that they interfere in the squared amplitude of the
exclusive photo-production process:
\begin{equation}
	|T|^2 = |T_{DVCS}|^2 + |T_{BH}|^2 + T_{DVCS}^* T_{BH} + T_{DVCS}T_{BH}^*. 
\end{equation}
The interference terms increase significantly the cross section in specific parts of the phase space
and lead to significant beam spin asymmetries (BSA), which are the focus of the measurements presented 
here.

Finally, we need to define our kinematics properly. We use the 
conventions from Fig. \ref{fig:PhiAngle} for angles and the experimental kinematic variables
used here are defined as: 
$-t = -(p_p-p_p^\prime)^2 = \Delta^2$ and 
$x_B = {Q^2 \over 2 M_N \nu}\sim {2\xi \over 1+\xi}$, with $M_N$ the nucleon mass and $\nu$ the
energy transfer to the target $\nu = {E -E^\prime} $.

\begin{figure}[tbp!]
\center
\includegraphics[width=9cm]{fig1/finalplane.png}
	\caption{Illustration of the scattering (or leptonic) and production (or hadronic) planes
	in the DVCS process.} 
\label{fig:PhiAngle}
\end{figure}

\subsection{Coherent nuclear DVCS}

The first reaction measured in the experiment is the coherent electro-production of a photon on helium
$e+^4\!\!He \rightarrow e+\gamma+^4\!\!He$ at high 4-momentum transfer squared ($Q^2$). The 
leading order diagram of the nuclear coherent DVCS is represented in Fig. \ref{fig:CohDiag}.
In our experiment, we focused on the measurement of the beam-spin asymmetry (BSA) noted $A_{LU}$ with 
$L$ for the longitudinally-polarized electron beam and $U$ the unpolarized target and defined as:  
\begin{equation}
A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
              {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
  \label{eq:BSA}
\end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the differential cross section for a positive 
(negative) beam helicity. At leading order and leading twist, the BSA can be expressed as \cite{Kirchner:2003wt}:        
\begin{eqnarray}
\label{eq:coh_BSA}
A_{LU}& =& \frac{x_A(1+\epsilon^2)^2}{y} \, s_1^{INT} \sin(\phi) \, 
\bigg/ \, \bigg[ \, \sum_{n=0}^{n=2}c_n^{BH}\cos{(n\phi)} +  \\
& & \frac{x_A^2 t {(1+\epsilon^2)}^2}{Q^2} P_1(\phi) P_2(\phi) \, c_0^{DVCS} + 
\frac{x_A (1+\epsilon^2)^2}{y} \sum_{n=0}^{n=1} c_n^{INT} \cos{(n\phi)} \bigg].  \nonumber 
\end{eqnarray}
where $\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ are the BH 
propagators. The factors: $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and 
$s_1^{INT}$ are the Fourier coefficients of the BH, the DVCS and the 
interference amplitudes for a spin-zero target. The explicit 
expressions of these coefficients, which have been derived based on the work of 
Kirchner and Müller \cite{Kirchner:2003wt}, can be found in Appendix \ref{app:Helium_cross_section}.

\begin{figure}[tbp!]
\center
\includegraphics[width=9cm]{fig3/DVCS_diagram.pdf}
\caption{Diagram representing the coherent nuclear DVCS.}
\label{fig:CohDiag}
\end{figure}

This formula can be expressed in a simplified manner for a spin-0 target as \cite{Belitsky:2008bz}:
\begin{equation}
A_{LU}(\phi) = \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the CFF $\mathcal{H}_{A}$ associated to the GPD $H_A$ of the 
spin-0 nucleus. The 
$\alpha_{i}$ factors are $\phi$-dependent kinematical terms that depend on the 
nuclear form factor $F_A$ and the independent variables $Q^2$, $x$ and $t$.  
These factors have the following simplified expressions:
\begin{eqnarray}
	\label{eq:alpha1}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
	\label{eq:alpha4}
\end{eqnarray}
where $S_{++}(1)$, $C_{++}(0)$, and $C_{++}(1)$ are the Fourier harmonics in the 
leptonic tensor \cite{Belitsky:2008bz}, their explicit expression can be found in 
appendix \ref{sec:eq}, and $x_{A} = \frac{M_{p}\cdot x}{M_{^4\!He}}$.

The Eq. \ref{eq:A_LU-coh} is particularly convenient to perform an extraction of 
the $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ through a fit of the BSA as 
a function of $\phi$. As can be seen in Fig. \ref{fig:alphas}, the form of each 
$\alpha$ coefficient has a characteristic $\phi$ dependence, such that a fit can
easily separate their respective contributions. The only caveat
is the large difference of magnitude between the $\alpha$ factors, which can lead to 
rather different error propagation for the two parts of the CFF.

\begin{figure}[tbp!]
\center
\includegraphics[width=8cm]{fig3/AlphaCoefs.png}
	\caption{Coefficients presented in Eq. \ref{eq:alpha1} to \ref{eq:alpha4}.
	Note the prescaling factors used for $\alpha_0$, $\alpha_2$ and $\alpha_3$.}
\label{fig:alphas}
\end{figure}

An important issue with the use of this theoretical framework is the large mass of the 
helium nuclei. Recent work indicate that the effect of this correction is moderate 
\cite{Braun:2012bg}, however the applicability to such a large mass remains to be
fully explored from the theoretical point of view.

\subsection{Incoherent nuclear DVCS}

The incoherent nuclear DVCS process, is the DVCS off a bound nucleon in a nucleus
as represented in Fig. \ref{fig:InCohDiag} for an helium-4 target. The remnants of 
the nucleus ($X$) contain only the missing three nucleons. 
The theory for incoherent DVCS on the nucleon is largely based on the free proton theory
already reviewed many times in the literature~\cite{Diehl:2003ny,Belitsky:2005qn,Guidal:2013rya}. 
Two important differences need to be accounted for however, the different initial state and the addition of 
final state effects. In the initial state, the intrinsic Fermi motion of the nucleons in the nucleus 
lead to an uncertainty on the exact kinematic of the reaction. Moreover, in general the nucleon 
is in an off-shell state that is not exactly identical to its final state. In the final state, 
interactions between the outgoing nucleon from the DVCS reaction and remnants of the nuclear 
target are possible. The latter leading to contamination from other channels, in particular 
charge exchange processes can lead to a large contribution from such background reaction.

\begin{figure}[tbp!]
\center
\includegraphics[width=10cm]{fig3/handbag_incoherent.pdf}
\caption{Diagram representing the incoherent nuclear DVCS.}
\label{fig:InCohDiag}
\end{figure}

Since, DVCS is a 
process selected using tight exclusivity constrains, some of the initial and final-state effects are
automatically mitigated. Selection criterion on missing energy and momentum are performed,
constraining the range of initial Fermi motion and final interactions possible. However, no theoretical
calculation is available to correct for the reminder of these effects yet. Since, modern calculations exist 
for such effects in DIS \cite{Cosyn:2017ekf}, we can expect them to be extended to the DVCS 
process as more data becomes available. Another avenue of progress
on this topic will be the use of experimental techniques like tagging. This process
can help to control both initial and final state effects by detecting the nuclear remnant.
In the tagged process the target breaks in two providing directly the initial state
of the struck nucleon, while a measured backward nuclear fragment limits significantly the
impact of final state effects.

\section{Past nuclear DVCS measurements}

The first measurement of nuclear DVCS has been performed by the HERMES 
collaboration~\cite{Airapetian:2009cga}. This experiment covered an array of 
nuclear targets and looked at the $A$ dependence of the
BSA signal. Their main results, reproduced in Fig. \ref{fig:HERMES1} and \ref{fig:HERMES2}, suffer
from large error bars, which makes them consistent with the free proton data and prevents us to
reach strong conclusions about possible nuclear effects. Yet, in the coherent DVCS case a rather
strong effect was expected, leading to a conflict between the HERMES results and theoretical
expectations.

\begin{figure}[tbp!]
\center
	\includegraphics[width=10.5cm]{fig3/HERMES_BSA.png}
	\caption{The sinus moment of the BSA as a function of $-t$ measured by HERMES
	for a series of nuclei \cite{Airapetian:2009cga}.}
\label{fig:HERMES1}
\end{figure}

\begin{figure}[tbp!]
\center
\includegraphics[width=10cm]{fig3/HERMES_BSA_2.png}
\caption{The sinus moment of the BSA at low and high $-t$ as a function of $A$ measured by HERMES
	\cite{Airapetian:2009cga}.}
\label{fig:HERMES2}
\end{figure}

An important issue with the HERMES measurement and how it is obtained from data has been raised in
\cite{Guzey:2003jh} to explain the discrepancy with theoretical expectations.
The main concern is that the DVCS process is not fully detected and the scattered target
is instead reconstructed through a missing mass measurement of the other reaction products. The 
issue with this method is that the detector resolution is not good enough to separate the 
coherent and incoherent channels properly. 
Instead, the results are labeled "coherent enriched" and "incoherent enriched" at low and high 
$-t$, respectively. This label is based on the assumption that the very different behavior of the
cross sections of the two channels in $t$ will lead to a clear differentiation. However, the
results in Fig. \ref{fig:HERMES2} show similar behaviors in both sectors of $t$, which is 
challenging this assumption and could explain the tension between theory and experiment.

Altogether, the large error bars and the impossibility to properly separate the coherent and 
incoherent channels have strongly impaired the measurement and any conclusions that could be
obtained from it. The CLAS experiment presented here has profited largely from this
result and was designed specifically to solve these two issues of low statistics and exclusivity.

\section{The CLAS nuclear DVCS experimental setup}

The CLAS experiment had for main objective to explore the coherent DVCS on helium-4 and assess if
the predicted BSA increase could be observed and if GPDs could be extracted from it. In order to
perform this measurement however, several instrumentation challenges needed to be resolved. First, to
measure the scattered electron and the low angle photon of the DVCS, we used CLAS in its 
DVCS setup, $i.e.$ with the addition of a low angle calorimeter and a 5-T solenoid magnet. Second, a
radial time projection chamber (RTPC) was installed to measure the
helium recoils and thus ensure the exclusivity of the process in the coherent channel. In this section, we 
will review the important elements of this detection setup.

\subsection{The CEBAF Large Acceptance Spectrometer (CLAS)}

The CLAS \cite{Mecking:2003zu} is installed in the hall-B of the JLab continuous electron beam 
accelerator facility (CEBAF). CLAS is specifically designed to study the multi-particles final states
that cannot be observed conveniently with arm spectrometers. Its historic use to measure DVCS in many different 
configurations made it an ideal place for this new DVCS measurement. CLAS is composed of six identical radial 
sectors separated by toroidal magnets, with each sector made of four detectors 
as shown in Fig. \ref{fig:CLAS}.  Three regions of drift chambers \cite{Mestayer:2000we} 
are placed in the solenoid area to reconstruct the charge particle's tracks and 
calculate their momentum. Then, an array of scintillators is placed to measure 
a precise time for each track and perform time-of-flight measurement 
\cite{Smith:1999ii}. These detectors cover the polar angle from 8 to 142 degrees. 
They are completed by two important forward detectors, covering the polar angle from 8 
to 45 degrees, a Cerenkov counter 
\cite{Adams:2001kk} and an electromagnetic calorimeter \cite{Amarian:2001zs}.

\begin{figure}[tbp!]
\center
\includegraphics[width=8cm]{fig3/CLAS_geantview-PS.jpg}
	\caption{View of the CLAS detector setup.}
\label{fig:CLAS}
\end{figure}

Altogether, CLAS provides a large acceptance of about $\pi$ steradians for momentums 
starting at 200 MeV/c. During the nuclear DVCS experiment,
the beam energy was maximal for CLAS with 6.064 GeV, while the beam intensity varied
between 120 and 150 nA. Such beam on our helium-4 target pressurized between 5 and 6 atm,
corresponds to luminosities in the range of $1$ to $1.2 \times 10^{34}$ cm$^{-2}$.s$^{-1}$.
During the experiment, the data acquisition rate ran around 3 kHz with about 70\% live-time
using an inclusive electron trigger.

\subsection{Adaptations for DVCS}

The CLAS collaboration has established a specific setup to measure the typically 
low angle photons of the DVCS process. This setup is composed of an inner 
calorimeter and a solenoid and has lead to numerous DVCS measurements on proton 
targets \cite{Seder:2014cdc,Jo:2015ema,HirlingerSaylor:2018bnu}.

The inner calorimeter, illustrated in Fig. \ref{fig:IC} is a homogeneous 
calorimeter composed 424 lead tungstate 
(PbWO) crystals read out by $5 \times 5$ mm$^2$ avalanche photo-diodes (APDs). 
It covers angles from 4 to 15 degrees. However, placing a detector at such low angles makes it 
particularly sensitive to the low energy Moller electrons scattered from the target.
To protect the calorimeter from this background, a 5~T solenoid is 
placed around the target to form
a magnetic shield. Thanks to this field, low energy charged particles (particularly 
electrons) curl around the beam line 
and never make it to the calorimeter or other CLAS detectors as illustrated 
by simulation results presented 
in Fig. \ref{fig:Solenoid}. This allows to run much higher luminosity experiments,
a necessity for low rates processes like DVCS.

\begin{figure}[tbp!]
\center
\includegraphics[width=8cm]{fig3/IC-CLAS.png}
	\caption{Representation of the inner calorimeter (IC) of CLAS. The crystals that compose the
	sensitive part of the detector are represented in purple.}
\label{fig:IC}
\end{figure}

\begin{figure}[tbp!]
\center
\includegraphics[width=14cm]{fig3/solenoid.png}
\caption{Representation of the center of CLAS with the beam background in red with and without
	the solenoid field activated, right and left, respectively.}
\label{fig:Solenoid}
\end{figure}

\subsection{The Radial time projection chamber}

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV/c at our beam energy. Therefore, a specific detector was 
needed to detect them. To design the present setup, large inspiration was drawn from the 
Bonus setup that also used a GEM based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
charges are projected toward large radii 
rather than towards the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}, here is a summary of key elements.

In order to detect the recoil helium nuclei from a DVCS reaction, we first need 
to ensure that it will come out of the target. For this, we used a light straw 
target made of a thin kapton wall of 27 $\mu$m filled with helium at 6 atm 
pressure. The entrance and exit windows are thin aluminum foils and an helium 
bag is placed downstream of the target to avoid interaction with air in the gap 
between the target and the beamline vacuum. The cylindrical 
chamber surrounds the target as illustrated in Fig.~\ref{fig:RTPCGlobal}, we 
list the elements composing it based on their radii:
\begin{itemize}
	\item Up to a radius of 3 mm the pressurized helium target.
	\item From 3 to 20 mm a keep-out zone filled with 1 atm of helium to 
		minimize the production of secondaries. 
	\item At 20 mm a grounded foil made of 4 $\mu$m aluminized Mylar to 
		isolate the chamber from the beam line 
		region and collect charges. It also serves to separate gas regions.
	\item From 20 to 30 mm a dead zone to separate the ground from the cathode
		filled with the drift gas, a mix of neon and dimethyl ether (DME) in 
		80/20 proportions.
	\item At 30 mm the cathode foil made of 4 $\mu$m aluminized Mylar.
	\item From 30 to 60 mm the drift region filled with the drift gas.
	\item From 60 to 69 mm the amplification regions, filled with drift gas, 
		with GEM foils placed at 60, 63 and 66 mm.
	\item At 69 mm the collection pads connected to the preamplifers placed 
		directly outside the chamber.
\end{itemize}

\begin{figure}[tbp!]
\center
\includegraphics[width=10cm]{fig3/RTPC_new_1.png}
\caption{Cut view of the RTPC.}
\label{fig:RTPCGlobal}
\end{figure}

The calibration of the detector has been performed with a 
dependence on $z$, the position along the beam line axis, due to 
variations in the magnetic fields. To perform this calibration we took  
dedicated data at 1.2 GeV beam energy. In 
this data set, we were able to select elastic events, for which the kinematic 
of the helium recoil can be calculated from the electron kinematic and 
directly compared to the measurement in the RTPC. This comparison helped to 
map the correspondence between time and position in the chamber and determine 
the drift path of electrons. A more detailed description of the calibration process 
is available in~\cite{Dupre:2017upj}.

\section{DVCS event selection}

\subsection{Particle identification}

The electrons are detected with the baseline CLAS detectors, the drift chamber measures the kinematic 
of the electron and the signal measured in both the Cerenkov counter and electromagnetic calorimeter
provide the identification. A signal of good quality is also required in the 
time-of-flight system, which serves as a time reference for other detectors. In particular, it 
serves for the identification of the protons. Several 
fiducial cuts are applied to ensure that particles did not go through part of the inner calorimeter 
or the solenoid, as well as to reject the edges of the detectors, where efficiency is rapidly 
decreasing. Kinematic corrections are also applied to the electrons and protons to correct for energy loss
and biases in calibration, they are at subpercent level except for protons below 500 MeV/c for which they
go up to 10\% at the detection limit of 200 MeV/c.

The photons from DVCS are mainly detected with the inner calorimeter. No specific 
identification cuts are used in this detector as large energy deposit are dominantly 
from electrons and photons, which cannot be separated reliably. However, the detection 
of an electron at larger angle in CLAS 
highly suppresses the number of electrons in the calorimeter, moreover
the exclusivity cuts used later in the analysis further this suppression. Left over accidentals, will be
accounted for in the background subtraction. The inner calorimeter was calibrated through a series
of steps, involving the reconstruction of $\pi^0$ from their two photons decay. Calibration was obtained 
with an iterative process to adjust each crystal gain to obtain the most accurate $\pi^0$ mass. A
global calibration of the calorimeter was also performed to account for incident angle, energy and 
time dependent effects.

Finally, we select events that contain a single electron, a high energy photon ($E>2$ GeV) and
either a helium or a proton. We apply a selection cut on the two charged particles 
to ensure they originate from the same vertex inside the target, thus rejecting accidentals and
events from the target windows.
Moreover, since we are aiming to study deep processes occurring at the partonic level, we select
$Q^2>1$ GeV$^2$. Also, the transferred momentum squared to the recoil $^{4}He$ is bound by
   a minimum value based on basic energy momentum conservation:
\begin{equation}
   t_{min} = - Q^{2} \frac{2(1- x_{A})(1 - \sqrt{1 + \epsilon ^{2}}) + \epsilon 
   ^{2}}{4 x_{A}(1-x_{A})+ \epsilon ^{2}},\\
\end{equation}
where $\epsilon ^{2} = \frac{4M^{2}_{^4\!He}x^{2}_{A}}{Q^{2}}$. For incoherent DVCS, 
we use a similar cut where $x_A$ is replaced by $x$ and $M_{^4\!He}$ by $M_{p}$.

\subsection{Exclusive photo-production selection}

To select exclusive events and suppress the backgrounds, we define several 
exclusivity cuts based on energy and momentum conservation. In principle, 
a selection based on two or three variables can be used to guaranty the 
exclusivity of the process. However, in such experiment, where particles 
are detected at very different energies and with very different detector 
resolutions, we chose to over constrain the selection by using seven 
variables. The seven variables are defined as follow for the coherent 
DVCS (replace helium by proton for the incoherent case):
\begin{itemize}
	\item Co-planarity ($\Delta \phi$) of the virtual photon, the real photon and
		the recoil helium, defined as the angle between the helium, virtual photon
		plane and the virtual photon, real photon plane;
	\item Missing energy of the full system;
	\item Missing mass of the full system;
	\item Missing transverse momentum of the full system;
	\item Missing mass of the system electron-helium, ignoring the photon;
	\item Missing mass of the system electron-photon, ignoring the recoil helium;
	\item Co-linearity ($\theta$) of the measured photon with the missing momentum of the 
		electron-helium system.
\end{itemize}

In the analysis, we apply selection cuts based on a fit of the exclusive peak at 3$\sigma$ around 
the mean value for each variable. This systematic method helps to avoid any bias 
in the selection of the events. The selection of coherent DVCS with these variables is illustrated in 
Fig. \ref{fig:CohExcCuts}. We note on these distributions only few minor anomalies, where the 
distribution have some asymmetries. These are linked with the detector resolutions, that impact
some of the kinematic variables non linearly. The selection of incoherent DVCS is presented in 
Fig. \ref{fig:IncExcCuts}, with two main differences: larger distributions overall and more
offset distributions. The former is largely due to Fermi motion, but simulations have shown
that this effect is not strong enough to reproduce these distribution widths and final 
state interactions must play a role as well. The later is caused by slight detector 
misalignment between CLAS sectors and are within the levels obtained with free proton 
targets~\cite{HirlingerSaylor:2018bnu} to which they can be directly compared.

\begin{figure}[tbp!]
\center
\includegraphics[trim=20 10 20 5,clip,width=15cm]{fig3/all_coh_exc_cuts.pdf}
	\caption{Distributions of the coherent photon production events before 
	(black) and after (blue line filled
	in Grey) the exclusivity cuts used to select coherent DVCS represented by the 
	red dashed lines. The histograms are shown as a function of the 
	seven variables used for the exclusivity selection described in the text, 
	plus the missing $P_x$ and $P_y$, in order left to right and top to bottom. 
	}
\label{fig:CohExcCuts}
\end{figure}

\begin{figure}[tbp!]
\center
\includegraphics[trim=20 10 20 5,clip,width=15cm]{fig3/all_incoh_exc_cuts.pdf}
\caption{Distributions of the incoherent photon production events before 
	(black) and after (blue line filled
	in Grey) the exclusivity cuts used to select coherent DVCS represented by the 
	red dashed lines. The histograms are shown as a function of the 
	seven variables used for the exclusivity selection described in the text, 
	plus the missing $P_x$ and $P_y$, in order left to right and top to bottom.}
\label{fig:IncExcCuts}
\end{figure}

\subsection{Background subtraction}

The main contamination is expected to be 
from the exclusive production of a $\pi^0$, the final state of which is very similar to DVCS with 
only an extra photon. In such an event, if one of the photon is produced at low energy, it is 
easy to confuse this process with single photon production. In order to estimate the contribution from
this channel in our data, we measure the exclusive $\pi^0$ production in the same way as DVCS, with 
a series of exclusivity cuts, completed by a selection cut on the invariant mass of the two photons
to match the $\pi^0$ mass. The events obtained for the coherent and incoherent
channels are shown in Fig. \ref{fig:CohPi0Simul} and \ref{fig:InCohPi0Simul}, respectively. From 
this sample, we elaborated an event generator, which after being processed
in the simulation of our detectors output the red histograms of Fig. \ref{fig:CohPi0Simul} and 
\ref{fig:InCohPi0Simul}. To correct the experimental data, we estimate the number of single photon 
events coming from DVCS with:
\begin{equation}
	N_{1\gamma,\pi^0}^{Exp} = \frac{N_{1\gamma,\pi^0}^{Sim}}{N_{2\gamma,\pi^0}^{Sim}} \times N_{2\gamma,\pi^0}^{Exp},\\
\end{equation}
where $N_{1\gamma,\pi^0}^{Sim}$ is the number of simulated exclusive $\pi^0$ mistaken for DVCS events,
$N_{2\gamma,\pi^0}^{Sim}$ the number of simulated exclusive $\pi^0$ fully reconstructed and $N_{2\gamma,\pi^0}^{Exp}$
the number of experimentally measured exclusive $\pi^0$. This number can then be subtracted to
the experimental measurement of DVCS events ($N_{DVCS}^{Exp}$) to get the corrected result: 
\begin{equation}
	N_{DVCS}^{Corr} = N_{DVCS}^{Exp} - N_{1\gamma,\pi^0}^{Exp}.\\
\end{equation}

\begin{figure}[p]
\center
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/xB_Coh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/Q2_Coh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/t_Coh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/phi_h_Coh_pi0.png}
	\caption{The measured (filled blue) and simulated (red) distributions of
	coherent exclusive $\pi^0$ production as a function of $x$, $Q^2$, $-t$ and $\phi$.}
\label{fig:CohPi0Simul}
\end{figure}

\begin{figure}[p]
\center
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/xB_InCoh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/Q2_InCoh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/t_InCoh_pi0.png}
\includegraphics[trim=70 15 70 70,clip,width=6.3cm]{fig3/pi0/phi_h_InCoh_pi0.png}
	\caption{The measured (filled blue) and simulated (red) distributions of
	incoherent exclusive $\pi^0$ production as a function of $x$, $Q^2$, $-t$ and $\phi$.}
\label{fig:InCohPi0Simul}
\end{figure}

The $\pi^0$ contamination was found to be 2 to 4\% in the coherent channel and 3 to 17\% in the incoherent
channel (with variations between bins). To make the correction on the DVCS BSA, we assume that
the exclusive $\pi^0$ production has no such asymmetry. This has been checked with the exclusive 
$\pi^0$ production data, for which no significant level of BSA has been measured.

The second important source of background comes from accidentals. Despite the many exclusivity cuts, it is 
possible to have particles from different events being combined and pass all the cuts to get 
into the data sample. To evaluate the number of such events, we invert the vertex selection 
of the two charged 
particles of the process, electron and helium (or proton in the incoherent case), and request that they
are separate. We find 4.1\% of the coherent and 6.5\% of the incoherent samples are accidentals.

\subsection{Systematic errors}

To evaluate the systematic error of the measurement, we performed several specialized studies. We 
evaluated the impact of changing the exclusivity selection cuts by varying them from 1 to 5 $\sigma$. 
We evaluated the impact of changing the binning in $\phi$ on the extraction of the beam-spin asymmetry
at 90°. The precision with which we know the beam polarization is estimated based on the spread 
of its measurements during the complete run period. We also checked different methods to make the 
simulation of the exclusive $\pi^0$ production to evaluate the 
possible bias introduced by this correction. As radiative corrections are expected to be small
for our process, we did not apply them, but associated an error of their expected size.
These errors are summarized in Tab. \ref{Table:systematic_uncertainties}, with their respective 
evaluated sizes. They are added quadratically for the total systematic error presented in the results.

\begin{table}[tbp]
\begin{center}
	\begin{tabular}{|m{4cm}|m{2cm}<{\centering}|m{2.3cm}<{\centering}|m{3.7cm}<{\centering}|}
\hline
\bf Systematic source & \bf  Coherent channel  & \bf Incoherent channel & \bf Type of systematic 
error\\
\hline
DVCS cuts & 8 $\%$ &  6 $\%$ & bin to bin\\
\hline
Data binning & 5.1$\%$ & 7.1$\%$ &bin to bin\\
\hline
Beam polarization &  3.5$\%$ &  3.5$\%$& Normalization\\
\hline
$\pi^0$ subtraction &  0.6$\%$ &  2.0$\%$ &bin to bin\\
\hline
%Beam charge asymmetry &  ~~~~~1.1$\%$&  ~~~~~1.1$\%$\\
%\hline
Radiative corrections &  0.1$\%$ & 0.1$\%$ & bin to bin\\
\hline
\textbf{Total} &  \textbf{10.1}$\%$ &   \textbf{10.1}$\%$ &bin to 
bin\\
\hline
\end{tabular}
\caption{The systematic uncertainties on the measured coherent and incoherent 
beam-spin asymmetries at $\phi = 90^{\circ}$.}
\label{Table:systematic_uncertainties}
\end{center}
\end{table}

As discussed above, the best way to define $t$ in the incoherent channel is not completely
straightforward. As can be seen in the Fig. \ref{fig:InCohDiag}, 
we can either use $t$ or $t^\prime$ ($= (p - p^\prime)^2$). In principle, the two are 
identical, but experimentally we face some issues. The measurement of $t$ is less precise than
$t^\prime$ because it involves the photon rather than charged particles. However, the 
exact measurement of $t^\prime$ is impossible and one needs to assume a proton at rest
in the initial state to calculate $t^\prime$. As it is not obvious which solution is best,
we studied the difference between the two results by analyzing the data independently using the two 
definitions. We found no significant difference between them, as is 
illustrated in Fig. \ref{fig:ttpComp}. We use in our final results $t$,
which is the correct definition with the slightly poorer resolution.

\begin{figure}[tbp!]
\center
\includegraphics[width=7.4cm]{fig3/ALU_90_p_vs_Q2_shortscenrario.pdf}
\includegraphics[width=7.4cm]{fig3/ALU_90_p_vs_t_shortscenrario.pdf}
%\includegraphics[width=7cm]{fig3/ALU_90_p_vs_x_shortscenrario.pdf}
	\caption{The BSA at 90° ($A_{LU}^{Incoh} (90^\circ)$) as a
	function of $Q^2$ and $-t$, using the photon based $t$ definition (empty red)
	and the proton based $t^\prime$ definition (full black).}
	%TODO we could remove HERMES and the models on these
\label{fig:ttpComp}
\end{figure}

\section{Results}

\subsection{Coherent DVCS}

In Fig. \ref{fig:CohALUphi}, we present the results for the BSA in the coherent DVCS channel. We 
observe the dominant sinusoidal component typical of the DVCS BSA, with an amplitude almost 
double than the one measured for the free proton \cite{Jo:2015ema}. This 
predicted feature of nuclear DVCS \cite{Guzey:2003jh} is observed here for the first time, we
can conclude that this measurement cleanly isolates the coherent DVCS process. The absence of 
this feature in the previous measurement from HERMES \cite{Airapetian:2009cga} and its clear observation here indicates that the recoil 
detection is a necessity to isolate the effects of the coherent DVCS process from backgrounds.

\begin{figure}[bp!]
\center
\includegraphics[width=12cm]{fig3/Coherent_ALU_phi.pdf}
	\caption{The BSA in the coherent exclusive photo-production off helium-4 as a 
	function of $\phi$ and $Q^2$ 
	(top panels), $x$ (middle panels) or $-t$ (lower panels). Error bars are  
	statistical, Grey bands represent the systematic errors. The full red lines show
	the fit of the data with the form of Eq.~\ref{eq:A_LU-coh}.}
	%TODO Add the chi2 somewhere
\label{fig:CohALUphi}
\end{figure}

\begin{figure}[tbp]
\center
\includegraphics[width=15cm]{fig3/Coherent_ALU_phi_90.pdf}
\caption{The BSA at 90° as a function of $Q^2$ (top panel), $x$ (middle panel) and $-t$ (lower panel).
	Our results are shown with black squares, HERMES results 
	with green circles \cite{Airapetian:2009cga} and theoretical predictions 
	from \cite{Liuti:2005gi} with full and dashed lines.}
\label{fig:CohALU90}
\end{figure}

We extract the BSA at 90° in Fig. \ref{fig:CohALU90}, the model compared to the 
data~\cite{Liuti:2005gi} is based on the hypothesis that the main nuclear effects are 
included by accounting for the nucleon off-shellness and kinematics in nuclei. It 
appears that this model undershoots our results systematically. However, a new independent 
calculation \cite{Fucini:2018gso}, using similar principles but with a more recent nuclear spectral 
function, has been able to reproduce the data very well. Since these results are from independent 
groups, it remains unclear at this point if the difference between them is manly due to the nuclear
spectral function or to its precise implementation. Another factor in the difference is that the
recent calculation from Fucini {\it et al.}~\cite{Fucini:2018gso} benefited from using the precise 
kinematic of each points. This feature appears to have a significant impact on some points, for 
instance the $-t$ distribution appears to have a peculiar structure well reproduced by the 
calculation. For this reason, we provide the detailed kinematic of each result presented in
appendix \ref{sec:fullresults}.

One of the promise of the helium-4 coherent DVCS was the ease to extract the CFF $\mathcal{H}_A$ from
data. To do so, we used the form from Eq. \ref{eq:A_LU-coh} to fit the data from Fig. 
\ref{fig:CohALUphi}. We obtained the real and imaginary parts of the unique helium-4
CFF presented in Fig. \ref{fig:CohCFF}. The results are rather encouraging, the two parts
of the CFF are constrained by data without need of any constraint beyond our data points. This 
capacity to obtain a
model independent result with such a limited data set offers a striking contrast with the
situation of the free proton fits \cite{Dupre:2016mai,Dupre:2017hfs}.

\begin{figure}[tbp!]
\center
\includegraphics[width=14cm]{fig3/Coherent_CFF.pdf}
	\caption{The imaginary (top panels) and real (bottom panels) parts of the helium-4
	CFF $\mathcal{H}_A$ as a function of $Q^2$ (left panels), $x$ (middle panels) and 
	$-t$ (right panels). The red full line red is the theoretical calculation from 
	Guzey {\it et al.}~\cite{Guzey:2003jh,Guzey:2008th}, the black dashed line is the same calculation 
	using the VGG model as input \cite{Vanderhaeghen:1999xj,Guidal:2004nd}, and
	the blue dashed line is a calculation from Liuti {\it et 
	al.}~\cite{Liuti:2005gi,GonzalezHernandez:2012jv}.} 
\label{fig:CohCFF}
\end{figure}

The CFF extraction allows us to compare the results to other theoretical calculations.
These are performed within the impulse approximation \cite{Guzey:2003jh,Guzey:2008th} 
and give the nuclear GPD directly from the proton and neutron GPDs. In 
Fig. \ref{fig:CohCFF}, we show two versions of this calculation, where two different nucleon GPD
models are used as input, compared with the calculation previously shown from Liuti 
{\it et al.}~\cite{Liuti:2005gi} with an updated nucleon model \cite{GonzalezHernandez:2012jv}.
We can see that the effect of changing the input nucleon GPD model
is of similar size or larger than the difference between nuclear 
models. However, at the level of precision of the present data,
it is not possible to resolve which variant is best. This feature highlights the 
importance of the choice of nucleon model to study nuclear effects with this data.

In summary, this measurement of the BSA in the deeply virtual coherent exclusive 
photo-production on a nucleus is the first to clearly isolate the effect of 
coherent nuclear DVCS and of nuclear GPDs. While, the statistical precision and 
the kinematic coverage are still far behind the experimental results on
proton target, the results appear to match very well the predictions using the GPD 
framework. Moreover, the extraction of the CFF appears to be very convenient based
on the BSA measurement only. Together, these findings  validate the relevance of the 
coherent nuclear DVCS to study the nucleus globally in terms of quarks and 
gluons.

\subsection{Incoherent DVCS}

\begin{figure}[bp!]
\center
\includegraphics[width=15cm]{fig3/incoherent_ALU_phi.pdf}
	\caption{The BSA in the incoherent exclusive photo-production off a proton bound in
	helium-4 as a function of $\phi$ and $Q^2$ 
	(top panels), $x$ (middle panels) or $-t$ (lower panels). Error bars are  
	statistical, Grey bands represent the systematic errors. The data is fitted with the 
	form $\frac{\alpha \sin(\phi)}{1+\beta \cos(\phi)}$; the results of the 
	fits are drawn with black full lines.}
\label{fig:InCohALUphi}
\end{figure}

The results for the measurement of the BSA in the incoherent DVCS channel are presented in
Fig. \ref{fig:InCohALUphi}. They display patterns rather similar to the one observed with the 
free proton, with a clear domination of their sinusoidal component. To compare the data to 
models, we extract the BSA at 90° with a fit of the form $\frac{\alpha \sin(\phi)}{1+\beta \cos(\phi)}$.

The asymmetries at 90° are presented in Fig. \ref{fig:IncALU} together with the theoretical
calculation presented also in Fig. \ref{fig:CohALU90}. The results are much more precise than 
the existing HERMES data and offer a strong constraint on the model presented. As in the
coherent case, the calculation appears to have issues to reproduce the shape of the data.
However, this time the calculation overshoots the data, sometimes by a significant amount.
An interesting way to look into this issue is to show the result on incoherent DVCS compared 
with the free proton. We can for instance make a ratio, in a fashion similar to the EMC 
effect, which allows to cancel out
the effects from the nucleon structure and highlight nuclear effects. Such ratio is presented
in Fig. \ref{fig:IncRatios}, where we observe again that most models significantly overshoot the data.

\begin{figure}[tbp!]
\center
\includegraphics[width=7.4cm]{fig3/ALU_90_p_vs_Q2_shortscenrario.png}
\includegraphics[width=7.4cm]{fig3/ALU_90_p_vs_t_shortscenrario.png}
\includegraphics[width=7.4cm]{fig3/ALU_90_p_vs_x_shortscenrario.png}
	\caption{The BSA at 90° as a function of  $Q^2$ (top left), $x$ (top right) and 
	$-t$ (bottom). Our measurement is represented with black squares, HERMES 
	measurement \cite{Airapetian:2009cga} with green circles and the theoretical 
	calculation from \cite{Liuti:2005gi} with full lines.}
\label{fig:IncALU}
\end{figure}

\begin{figure}[tbp!]
\center
\includegraphics[width=7.4cm]{fig3/ALU_ratioInc_Q2_shortscenrario.png}
\includegraphics[width=7.4cm]{fig3/ALU_ratioInc_t_shortscenrario.png}
\includegraphics[width=7.4cm]{fig3/ALU_ratioInc_x_shortscenrario.png}
	\caption{DVCS BSA ratio of the bound proton to the free proton as a function of 
	$Q^2$ (top left), $x$ (top right) and $-t$ (bottom). The present measurement is 
	represented with black squares, HERMES 
	measurement \cite{Airapetian:2009cga} with green circles, the theoretical 
	calculation from \cite{Liuti:2005gi} with blue and red full lines and the
	calculation from \cite{Guzey:2008fe} with full and dashed black lines.}
\label{fig:IncRatios}
\end{figure}

This measurement of the incoherent DVCS BSA shows a global tendency of the experimental data 
to be smaller than expected from theoretical calculations. We observe the incoherent DVCS 
BSA to go as low as 60\% of the size of the BSA observed
on the free proton at $x_B \sim 0.15$. This drastic reduction is still unexplained, though some theoretical 
work has been recently published to prolong the calculation from \cite{Fucini:2018gso} to the incoherent 
channel \cite{Fucini:2019xlc}. The explanation for this surprising behavior can come from different sources
both in the initial state and the final state.

Further work is needed to fully comprehend this newly discovered nuclear effect. On the 
experimental side, the use of tagging methods, where the nuclear fragments are measured
appear to offer the best option forward. Indeed, tagging offers the best chance to understand better
this result by offering better control over both the initial and the final state 
effects~\cite{Dupre:2015jha}.

\section{Summary}

We report the measurement of the coherent and incoherent DVCS processes off helium-4 with
CLAS at JLab. To properly isolate the coherent channel, the experiment used a specially 
designed RTPC to detect the scattered helium-4. While statistically limited, this coherent 
DVCS measurement reveals the large BSA ($\sim 35$\%) expected by theoretical calculations
made in the impulse approximation. Moreover, we showed that the CFF extraction can be 
immediately performed using these data without any model prescription. The incoherent DVCS measurement 
however reveals relatively small asymmetries in comparison to previous free proton 
measurements. It remains unclear what is the source of the suppression of the BSA as both 
initial and final state effect contributions could lead to such outcome.

% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}

% Surround figure environment with turnpage environment for landscape
% figure
% \begin{turnpage}
% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}
% \end{turnpage}

% tables should appear as floats within the text
%
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
% The ruledtabular enviroment adds doubled rules to table and sets a
% reasonable default table settings.
% Use the table* environment to get a full-width table in two-column
% Add \usepackage{longtable} and the longtable (or longtable*}
% environment for nicely formatted long tables. Or use the the [H]
% placement option to break a long table (with less control than 
% in longtable).
% \begin{table}%[H] add [H] placement to break table across pages
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% Lines of table here ending with \\
% \end{tabular}
% \end{ruledtabular}
% \end{table}

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
%\appendix
%\section{}

% Surround table environment with turnpage environment for landscape
% table
% \begin{turnpage}
% \begin{table}
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% \end{tabular}
% \end{ruledtabular}
% \end{table}
% \end{turnpage}

\begin{acknowledgments}
	%TODO Add Sergio and ERC!
 The authors acknowledge the staff of the Accelerator and Physics Divisions at 
the Thomas Jefferson National Accelerator Facility who made this experiment 
possible. This work was supported in part by the Chilean Comisi\'on Nacional de 
Investigaci\'on Cient\'ifica y Tecnol\'ogica (CONICYT), by CONICYT PIA grant 
ACT1413, the Italian Instituto Nazionale di Fisica Nucleare, the French Centre 
National de la Recherche Scientifique, the French Commissariat \`a l'Energie 
Atomique, the U.S. Department of Energy under Contract No. DE-AC02-06CH11357, 
the United Kingdom Science and Technology Facilities Council (STFC), the 
Scottish Universities Physics Alliance (SUPA), the National Research Foundation 
of Korea, and the Office of Research and Economic Development at Mississippi 
State University.  M.~Hattawy also acknowledges the support of the Consulat 
G\'en\'eral de France \`a J\'erusalem.  The Southeastern Universities Research 
Association operates the Thomas Jefferson National Accelerator Facility for the 
United States Department of Energy under Contract No. DE-AC05-06OR23177.
\end{acknowledgments}

\appendix
\section{Expressions for the BSA of the coherent DVCS}
\label{sec:eq}

We present in this appendix the detailed expressions used for Eq. \ref{eq:coh_BSA} 
and Eq. \ref{eq:alpha1} to \ref{eq:alpha4}. These are adapted from the work of 
Kirchner and Müller \cite{Kirchner:2003wt} to match the notations and conventions used
in this work.

First, $\mathcal{P}_{1}(\phi)$ and $\mathcal{P}_{2}(\phi)$ are BH propagators and defined as:
\begin{align}
&\mathcal{P}_{1}(\phi) = \frac{(k - q')^{2}}{Q^{2}} = - \frac{1}{y (1 + \epsilon^{2})} 
\big[ J + 2 K \cos(\phi) \big] \\
&\mathcal{P}_{2}(\phi) = \frac{(k - \Delta)^{2}}{Q^{2}} = 1 + \frac{t}{Q^{2}} + 
\frac{1}{y (1 + \epsilon^{2})} \big[ J + 2 K \cos(\phi) \big]
\end{align}
with,
\begin{align}
& J = \bigg( 1 - y - \frac{y \epsilon^{2}}{2} \bigg) \bigg(1 + \frac{t}{Q^{2}} \bigg) - 
(1 - x_{A})(2 - y) \frac{t}{Q^{2}} \\
& K^{2} = - \delta t \, (1 - x_{A}) \bigg( 1 - y - \frac{y^{2} \epsilon^{2}}{4} \bigg) 
\bigg\{ \sqrt{1 + \epsilon^{2}} + \frac{4 x_{A} (1-x_{A}) + \epsilon^{2}}{4 (1 - x_{A})}
\delta t \bigg\} \\
& \delta t = \frac{t - t_{min}}{Q^{2}} = \frac{t}{Q^{2}} + \frac{2(1-x_{A}) \left(1- \sqrt{1 + 
\epsilon^{2}} \right) + \epsilon^{2}}{4 x_{A} (1- x_{A}) + \epsilon^{2}}.
\end{align}

The Fourier coefficients for BH contributions are defined as:
\begin{eqnarray}
c_0^{BH} = & \bigg[ & \left\{ {(2-y)}^2 + y^2{(1+\epsilon^2)}^2 \right\} 
\left\{ \frac{\epsilon^2 Q^2}{t} + 4 (1-x_A) + (4x_A+\epsilon^2) \frac{t}{Q^2} 
\right\} \nonumber \\
& \phantom{\bigg[} & + 2 \epsilon^2 \left\{ 4(1-y)(3+2\epsilon^2) + y^2(2-\epsilon^4) 
\right\} - 4 x_A^2{(2-y)}^2 (2+\epsilon^2) \frac{t}{Q^2} \nonumber \\
& \phantom{\bigg[} & + 8 K^2 \frac{\epsilon^2 Q^2}{t} \,\,\,\,\,\,\, \bigg] F_A^2(t)  \\
c_1^{BH} = & \phantom{\bigg[} & -8 (2-y) K \left\{ 2 x_A + \epsilon^2 - 
\frac{\epsilon^2 Q^2}{t} \right\} F_A^2(t)  \\
c_2^{BH} = & \phantom{\bigg[} & 8 K^2 \frac{\epsilon^2 Q^2}{t} F_A^2(t) 
\end{eqnarray} 
where $F_A(t)$ is the electromagnetic form factor of the $^4$He. 
The coefficient for the DVCS contribution is given by: 
\begin{equation}
   c_0^{DVCS}= 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + \epsilon^2} \, 
   {\mathcal H}_A {\mathcal H}^{\star}_A 
   \label{eq:c0DVCS}
\end{equation}

Finally, the interference amplitude coefficients are written as:
\begin{equation}
s_{1}^{INT} = F_{A}(t) \Im m(\mathcal{H}_{A}) S_{++}(1),
\end{equation}
with
\begin{eqnarray}
   S_{++}(1) &=& \frac{-8K(2-y)y}{1+\epsilon^2} \left( 1 + 
\frac{1-xA+\frac{\sqrt{1+\epsilon^2}-1}{2}}{1+\epsilon^2} 
\frac{t-t_{min}}{Q^{2}} \right) \label{eq:s1I}
\end{eqnarray}

\begin{eqnarray}
c_0^{INT} &=& F_A(t) \Re e(\mathcal{H}_{A}) C_{++}(0),
\end{eqnarray}
with \begin{eqnarray}  C_{++}(0) &=&
\frac{-4(2-y)(1+\sqrt{1+\epsilon^{2}})}{(1+\epsilon^{2})^2}  \bigg\{ 
   \frac{\widetilde{K}^2}{Q^2}  \frac{(2-y)^2}{\sqrt{1+\epsilon^{2}}} \, \\
   &+& \frac{t}{Q^2}  \left( 1 - y - \frac{\epsilon^2}{4} y^2 \right)  
(2-x_{A}) \left(  1 + \frac{2x_A(2-x_A + \frac{\sqrt{1+\epsilon^{2}}-1}{2} + 
\frac{\epsilon^{2}}{2x_A})\frac{t}{Q^2} + \epsilon^{2}}{(2-x_A) 
(1+\sqrt{1+\epsilon^{2}})}  \right)  \bigg\} \nonumber
 \label{eq:c0I} 
 \end{eqnarray}

\begin{eqnarray}
   c_1^{INT} &=&  F_A(t) \Re e(\mathcal{H}_{A}) C_{++}(1),
\end{eqnarray}
with  
   \begin{eqnarray}
   C_{++}(1) &=&
   \frac{-16K(1-y+\frac{\epsilon^{2}}{4}y^2)}{(1+\epsilon^{2})^{5/2}}\bigg\{\left(1+(1-x_A)\frac{\sqrt{1+\epsilon^{2}}-1}{2x_A} 
   + \frac{\epsilon^{2}}{4x_A}\right) 
\frac{x_At}{Q^2}-\frac{3\epsilon^{2}}{4.0} \bigg\} \nonumber \\&-& 4K \left( 
2-2y+y^2+\frac{\epsilon^{2}}{2}y^2\right)\frac{1+\sqrt{1+\epsilon^{2}}-\epsilon^{2}}{(1+e2)^{5/2}}\bigg\{1-(1-3x_A)\frac{t}{Q^2}\nonumber\\&\,\,\,\,&\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+\frac{1-\sqrt{1+\epsilon^{2}}+3\epsilon^{2}}{1+\sqrt{1+\epsilon^{2}}-\epsilon^{2}} 
\frac{x_A*t}{Q^2}\bigg\} \label{eq:c1I}
\end{eqnarray}

\section{Table of results with information on full kinematics}
\label{sec:fullresults}
 tests

% Create the reference section using BibTeX:
\bibliography{hadro.bib}

\end{document}